### 2015-03-22 inst/examples/Vaasa.Rmd
Vaasa, osa 7

---
title: "Julkaisutoiminta Vaasassa 1764-1917"
author: "Niko Ilomäki"
date: "`r Sys.Date()`"
output: markdown_document
---

```{r, echo=FALSE, message=FALSE}
library(dplyr)
```

```{r, echo=FALSE, message=FALSE}
# data ei välttämättä lopullinen
df <- readRDS("df.Rdata")
df <- df %>% filter(!is.na(published_in))
vaasa1 <- df %>% filter(city == "Vaasa") %>% group_by(published_in) %>% tally() %>% arrange(published_in)
vaasa1$decade <- as.numeric(paste0(substr(vaasa1$published_in,1,3),"0"))
vaasa2 <- vaasa1 %>% group_by(decade) %>% summarise(m = sum(n)/10)
vaasa <- inner_join(vaasa1,vaasa2)
```

```{r, echo=FALSE, fig.width=20, fig.height=10}
par(las=1)
par(mgp=c(10,3,0))
par(mar=c(7,17,6,4))
plot(vaasa$published_in,vaasa$n,cex=1.8,pch="|",cex.main=5.0,main="Julkaisutoiminta Vaasassa 1764-1917",cex.lab=5.0,xlab="",ylab="Dokumenttia / vuosi",axes=FALSE,col="black")
lines(vaasa$published_in,vaasa$m,type="l",lty=3,col="black",lwd=4)
axis(side = 1, at = c(1764,1809,1863,1917),cex.axis=3.2)
axis(side = 2, at = c(0,40,80,120,160,194),cex.axis=3.2)
```

```{r, echo=FALSE, fig.width=20, fig.height=10}
par(las=1)
par(mgp=c(10,3,0))
par(mar=c(7,17,6,4))
plot(vaasa$published_in,vaasa$n,cex=1.8,pch="|",cex.main=5.0,main="Julkaisutoiminta Vaasassa 1764-1917",cex.lab=5.0,xlab="",ylab="Dokumenttia / vuosi",axes=FALSE,col="black")
lines(vaasa$published_in,vaasa$m,type="l",lty=3,col="black",lwd=4)
axis(side = 1, at = c(1764,1800,1840,1880,1917),cex.axis=3.2)
axis(side = 2, at = c(0,40,80,120,160,194),cex.axis=3.2)
```