### 2015-05-10 inst/examples/suvut.Rmd
Rmd fixes

---
title: "Suvut"
author: "Niko Ilomäki"
date: "`r Sys.Date()`"
output: markdown_document
---

```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(stringr)
```

```{r suvut, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=15}
df <- readRDS("df.Rdata")
df <- df %>% filter(!is.na(family_name))
df2 <- df %>% group_by(family_name) %>% summarise(members = n_distinct(first_name))
df2 <- df2 %>% filter(members > 10) %>% arrange(desc(members))
theme_set(theme_bw(15))
df2$extra <- reorder(df2$family_name, df2$members)
p <- ggplot(df2, aes(x=family_name, y=members))
p <- p + geom_bar(aes(x=extra), data=df2, stat = "identity")
p <- p + coord_flip()
p <- p + ylab("")
p <- p + xlab("")
p <- p + ggtitle("Etunimiä per sukunimi")
print(p)
```