### 2015-12-07 inst/examples/Paper.Rmd
Some plots

---
title: "Paper Consumption in Finland, 1488-1828"
author: "Niko Ilomäki"
date: "December 7, 2015"
output: markdown_document
---

```{r summaryinit, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
library(stringr)
library(dplyr)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)
df <- readRDS("df.Rds")
dfs <- df %>% filter(publication_year > 1400 & publication_year < 1829) %>%
  filter(!is.na(paper.consumption.km2) & paper.consumption.km2 > -0.01) %>%
  group_by(publication_year) %>%
  summarise(titles = n(), paper = sum(paper.consumption.km2))
authors <- df %>% filter(publication_year > 1400 & publication_year < 1829) %>%
  group_by(author_name) %>%
  summarise(titles=n()) %>% arrange(desc(titles))
top10 <- tail(head(authors,11),10)
dfs10 <- df %>% filter(publication_year > 1400 & publication_year < 1829) %>%
  filter(author_name %in% c(top10$author_name)) %>%
  filter(!is.na(paper.consumption.km2) & paper.consumption.km2 > -0.01) %>%
  group_by(publication_year) %>%
  summarise(titles = n(), paper = sum(paper.consumption.km2))
```

```{r titles, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=4, fig.width=8}
p1 <- ggplot(dfs, aes(y = titles, x = publication_year))
p1 <- p1 + geom_smooth(color = "black") + geom_point()
p1 <- p1 + xlab("Publication Year") + ylab("Title Count") 
p1 <- p1 + theme(axis.text.x = element_text(size = 20), axis.text.y = element_text(size = 20))
p1 <- p1 + theme(axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20))
p1 <- p1 + guides(fill = "none", alpha = "none")
p1 <- p1 + ggtitle("Finnish publications 1488-1828: title count (all authors)")
print(p1)
```

```{r paper, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=4, fig.width=8.2}
p2 <- ggplot(dfs, aes(y = paper, x = publication_year))
p2 <- p2 + geom_smooth(color = "black") + geom_point() +
       xlab("Publication Year") +
       ylab("Paper Consumption (km2)")

p2 <- p2 + theme(axis.text.x = element_text(size = 20), axis.text.y = element_text(size = 20))
p2 <- p2 + theme(axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20))
p2 <- p2 + guides(fill = "none", alpha = "none")
p2 <- p2 + ggtitle("Finnish publications 1488-1828: paper consumption (all authors)")
print(p2)
```

```{r titles10, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=4, fig.width=8}
p1b <- ggplot(dfs10, aes(y = titles, x = publication_year))
p1b <- p1b + geom_smooth(color = "black") + geom_point()
p1b <- p1b + xlab("Publication Year") + ylab("Title Count") 
p1b <- p1b + theme(axis.text.x = element_text(size = 20), axis.text.y = element_text(size = 20))
p1b <- p1b + theme(axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20))
p1b <- p1b + guides(fill = "none", alpha = "none")
p1b <- p1b + ggtitle("Finnish publications 1488-1828: title count (top 10 authors)")
print(p1b)
```

```{r paper10, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=4, fig.width=8.2}
p2b <- ggplot(dfs10, aes(y = paper, x = publication_year))
p2b <- p2b + geom_smooth(color = "black") + geom_point() +
       xlab("Publication Year") +
       ylab("Paper Consumption (km2)")

p2b <- p2b + theme(axis.text.x = element_text(size = 20), axis.text.y = element_text(size = 20))
p2b <- p2b + theme(axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20))
p2b <- p2b + guides(fill = "none", alpha = "none")
p2b <- p2b + ggtitle("Finnish publications 1488-1828: paper consumption (top 10 authors)")
print(p2b)
```