### 2015-12-08 inst/examples/20151209-HelsinkiDH.Rmd
20151209 pics

---
title: "Digital History in Finland, 1488-1828"
author: "Leo Lahti"
date: "December 9, 2015"
output: markdown_document
---

This markdown document contains presentation material for [Digital History in Finland seminar (University of Helsinki, Dec 9, 2015)](http://blogs.helsinki.fi/mstolone/2015/09/21/call-for-papers-digital-history-in-finland-wednesday-9-12-2015/). The source code to reproduce the figures is in the [Rmarkdown document](https://github.com/rOpenGov/fennica/blob/master/inst/examples/20151209-HelsinkiDH.Rmd).

```{r 20151209-HDH-init, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
library(stringr)
library(dplyr)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)

# Nice theme
theme_set(theme_bw(20))

# Read data
dfo <- readRDS("df.Rds")

# Add publication decade
# TODO add already in the original preprocessed data matrix
dfo$publication_decade <- 10*floor(dfo$publication_year/10)

# Remove very early and late years
dfo <- dfo %>% filter(publication_year > 1400 & publication_year < 1911)
```

```{r 20151209-HDH-TurkuVSOthers, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=4, fig.width=8}

# Let us focus on years 1400-1828
df <- dfo %>% filter(publication_year > 1400 & publication_year < 1829)

# Add indicator field: Turku or Not Turku
myplace <- "Turku"
df$myplace <- rep(NA, nrow(df))
df$myplace[df$publication_place == myplace] <- myplace
df$myplace[!df$publication_place == myplace & df$country == "Finland"] <- "Finland"
df$myplace[!df$country == "Finland"] <- "Foreign"
df$myplace <- factor(df$myplace, levels = rev(c("Foreign", "Finland", "Turku")))

# Keep entries with known publication place and paper consumption
# Group by year/decade and Turku/Other
dfs <- df %>%
  #filter(publication_year >= 1600) %>%
  filter(!is.na(publication_place) & !is.na(paper.consumption.km2) & paper.consumption.km2 > -0.01) %>%
  group_by(publication_decade, myplace) %>%
  summarise(titles = n(), paper = sum(paper.consumption.km2))

p <- ggplot(dfs, aes(x = publication_decade, y = paper, fill = myplace, order = -as.numeric(myplace))) +
     geom_bar(position = "stack", stat = "identity") +
     ggtitle("Paper consumption") +
     scale_fill_manual(values = c("blue", "black", "darkgreen")) +
     xlab("Publication decade") + ylab("Paper (km2)")
print(p)

p <- ggplot(dfs, aes(x = publication_decade, y = titles, fill = myplace, order = -as.numeric(myplace))) +
     geom_bar(position = "stack", stat = "identity") +
     ggtitle("Title count") +
     scale_fill_manual(values = c("blue", "black", "darkgreen")) +
     xlab("Publication decade") + ylab("Title count (n)")
print(p)

#p + guides(fill = guide_legend(order = 1))
```

```{r 20151209-HDH-Author, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=4, fig.width=6}
p1 <- top_plot(dfo, "author", 10) + ggtitle(paste("Top authors (", min(na.omit(dfo$publication_year)), "-", max(na.omit(dfo$publication_year)), ")", sep = "")) + ylab("Title count (n)")
print(p1)

p2 <- top_plot(df, "author", 10) + ggtitle(paste("Top authors (", min(na.omit(df$publication_year)), "-", max(na.omit(df$publication_year)), ")", sep = "")) + ylab("Title count (n)")
print(p2)
```