### 2016-04-02 inst/examples/summary.Rmd
ok

---
title: "Preprocessing summary"
author: "Leo Lahti and Niko Ilomäki"
date: "March 2, 2016"
---

# Fennica summary

## Annotated documents

Fraction of documents with entries for each annotation field. For a full list of summaries for field contents, see [here](https://github.com/rOpenGov/fennica/tree/master/inst/examples/output.tables).

```{r missing, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=15}
library(stringr)
library(dplyr)
library(reshape2)
library(ggplot2)
ntop <- 20

# Read the preprocessed data
df <- readRDS("df.Rds")
df.orig <- readRDS("df.orig.Rds")
conversions <- readRDS("conversions.Rds")
```

## Publisher

[Self-published accepted](https://github.com/rOpenGov/fennica/blob/master/inst/examples/output.tables/self_published_accepted.csv) (Check for formatting and synonymes)

[Self-published raw fields](https://github.com/rOpenGov/fennica/blob/master/inst/examples/output.tables/self_published_rawfields.csv) (Check that the self-publication is a correct classification for these docs)

We also have corporate info available ?

[Accepted corporate info](https://github.com/rOpenGov/fennica/blob/master/inst/examples/output.tables/corporate_accepted.csv)

[Discarded corporate info](https://github.com/rOpenGov/fennica/blob/master/inst/examples/output.tables/corporate_discarded.csv)

## Subject geographical places

```{r geo, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
tab <- rev(sort(table(unlist(strsplit(as.character(df$subject_geography), ";")))))
```

Top-`r ntop` geographical places are shown together with the number of documents assigned to that region. Geography information is available for `r sum(!is.na(df$subject_geography))` documents (`r round(100*mean(!is.na(df$subject_geography)))`%). There are `r length(tab)` unique geographical places.

```{r geo2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
ntop <- 50
par(mar = c(5, 10, 3, 1)); barplot(log10(rev(tab[1:ntop])), horiz = T, las = 1, cex.names = 0.8, xlab = "Documents (log10)", main = "Most common geographic places")
```

```{r geo3, echo=FALSE, message=FALSE, warning=FALSE}
f <- "output.tables/geoplaces.csv"
write.table(cbind(Geography = names(tab), Documents = tab), file = f, sep = "|", quote = F, row.names = F)
```

[Complete geoplace listing](https://github.com/rOpenGov/fennica/blob/master/inst/examples/output.tables/geoplaces.csv)