### 2016-04-02 inst/extras/misc.Rmd
ok

## Topics

```{r topics1, topics, echo=FALSE, message=FALSE, warning=FALSE}
# List all topics
spl <- strsplit(na.omit(as.character(df$topic)), ";")

# Topics per document
# hist(sapply(spl, length))

# Documents per topic
tab <- sort(table(unlist(spl)))
tab <- tab[!names(tab) == "NA"]
tab <- rev(sort(tab)) 
```

```{r topics2, echo=FALSE, message=FALSE, warning=FALSE}
# Write to file
f <- "output.tables/subjecttopics.tab"
write.table(cbind(Topic = sort(names(tab))), file = f, sep = "\t", quote = F, row.names = F)
ntop <- 50
```

[Complete subject topic counts](https://github.com/rOpenGov/fennica/blob/master/inst/examples/output.tables/subjecttopics.tab)

Top-`r ntop` topics and number of documents for each. In total, there are `r length(unique(df$topic))` unique topics and `r sum(!is.na(df$topic))` documents assigned to one or more topics (`r round(100*mean(!is.na(df$topic)))`).

```{r topics3, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
par(mar = c(5, 10, 3, 1)); barplot(rev(rev(sort(table(df$topic)))[1:ntop]), horiz = T, las = 1, cex.names = 0.8, xlab = "Documents", main = "Most common topics")
```

Conversions from the original to preprocessed fields:

```{r summary-conversions, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=15}
a <- harmonize_field_names()
b <- sapply(conversions, function (x) {paste(x, collapse = ";")})
coms <- intersect(as.character(a$name), names(b))

field.conversions <- cbind(a[match(coms, a$name), ], preprocessed = unname(b[coms]))
names(field.conversions) <- c("marc_clearname", "marc_field", "preprocessed_fields")
field.conversions <- field.conversions[, c(2, 1, 3)]
kable(field.conversions)
```

Percentage of non-NA entries among all documents in the preprocessed data (note that in some cases a significant fraction of this information is missing already in the raw data)

```{r missingb, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=15}
# Need to use this trick to have NAs for all data types
missing <- 100*apply(df, 2, function (x) {mean(is.na(gsub("^NA$", NA, as.character(x))))})
df2 <- data.frame(list(missing = missing, field = names(missing)))
df2$field <- factor(df2$field, levels = df2$field[rev(order(df2$missing))])

theme_set(theme_bw(15))
p <- ggplot(df2, aes(x = field, y = 100 - missing))
p <- p + geom_bar(stat = "identity")
p <- p + coord_flip()
p <- p + ylab("")
p <- p + xlab("")
p <- p + ggtitle("Documents with data (%)")
print(p)
```

## Publication time

[Successfully converted publication times](output.tables/publication-time-accepted.csv)

[Discarded publication times](output.tables/publication-time-discarded.csv)

```{r publication_time, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
#hist(df$publication_year, main = "Document frequencies over time")
dfs <- df %>% group_by(publication_year) %>% tally()
ggplot(dfs, aes(x = publication_year, y = n)) + geom_point() + ggtitle("Document frequencies over time") + ylab("Title count") + xlab("Publication year")
```

## Publication places

Top-`r ntop` publication places are shown together with the number of documents. This info is available for `r sum(!is.na(df$publication_place))` documents (`r round(100*mean(!is.na(df$publication_place)))`%). There are `r length(unique(str_trim(unlist(strsplit(as.character(df$publication_place), ";")))))` unique publication places.

```{r publication_place, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
tab <- rev(sort(table(str_trim(unlist(strsplit(as.character(df$publication_place), ";"))))))
ntop <- 50
par(mar = c(5, 10, 3, 1)); barplot(log10(rev(tab[1:ntop])), horiz = T, las = 1, cex.names = 0.8, xlab = "Documents (log10)", main = "Most common publication places")
```

[Successfully converted publication places](https://github.com/rOpenGov/fennica/blob/master/inst/examples/output.tables/publication_place_accepted.csv)

[Discarded publication places](https://github.com/rOpenGov/fennica/blob/master/inst/examples/output.tables/publication_place_discarded.csv)

[Publication places missing country information](https://github.com/rOpenGov/fennica/blob/master/inst/examples/output.tables/publication_place_missingcountry.csv)

## Dimensions

[Successfully converted dimension info](https://github.com/rOpenGov/fennica/blob/master/inst/examples/output.tables/accepted_dimensions.csv)

[Discarded dimension info](https://github.com/rOpenGov/fennica/blob/master/inst/examples/output.tables/missing_dimensions.csv)

[Discarded gatherings info](https://github.com/rOpenGov/fennica/blob/master/inst/examples/output.tables/missing_gatherings.csv)

## Authors

[Accepted unique authors](https://github.com/rOpenGov/fennica/blob/master/inst/examples/output.tables/author_accepted.csv)

[Discarded authors](https://github.com/rOpenGov/fennica/blob/master/inst/examples/output.tables/author_discarded.csv)

[Discarded author first names](output.tables/author_name_discarded_first.csv)

[Discarded author last names](output.tables/author_name_discarded_last.csv)

[Authors missing life years](output.tables/authors_missing_lifeyears.csv)

Top-`r ntop` uniquely identified authors and number of documents for each (duplicate docs not checked yet).

```{r authors, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
a <- rev(rev(sort(table(df$author)))[2:ntop+1])
par(mar = c(5, 10, 3, 1)); barplot(a, horiz = T, las = 1, cex.names = 0.8, xlab = "Documents", main = "Top authors")
```