### 2016-07-12 inst/examples/20160715-slides-Krakow.Rmd
sync

---
title: "Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828"
author: "Mikko Tolonen, Jani Marjanen, Hege Roivainen, Leo Lahti"
date: "`r Sys.Date()`"
output: 
  beamer_presentation:
    theme: "default"
    colortheme: "dolphin"
---

```{r 201606krakow-init, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Default time span
min.year <- 1488
max.year <- 1828

library(ggmap)
library(gisfin)
library(stringr)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)
library(devtools)
# install_github("ropengov/gisfin")
# devtools::install_github('cttobin/ggthemr')
library(gisfin)
#library(ggthemr)

knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(fig.path = "slides_201606_Krakow/", dev="CairoPNG")
knitr::opts_chunk$set(fig.path = "slides_201606_Krakow/")

# Set locale
tmp <- Sys.setlocale(locale="UTF-8")

# Nice theme
theme_set(theme_bw(26))

# Nice default themes
# https://github.com/cttobin/ggthemr#palettes
#ggthemr('fresh', text_size = 20)
# ggthemr('greyscale', text_size = 20)
#ggthemr('light', text_size = 20)
# ggthemr('pale', text_size = 20)
#ggthemr_reset() # Reset theme
```

```{r 201606krakow-init2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Full combined catalogue (Fen + Kun) with marked duplicates
df.combined.preprocessed <- readRDS("df.combined.Rds")
# Data with duplicates removed
df.full <- df.combined.preprocessed %>% filter(!remove)
# Data with years limited as well
df0 <- df.full %>% filter(publication_year >= min.year & publication_year <= max.year)
```

---

### TODO

sivu 12. Gävle-kuva vaihdettu Tukholmaan, pitää tsiigata mitkä vuodet highlightataan. Tällaisena ei vielä toimi. Onhan katalogi oikein?

sivu 27. Dissertation languages: kategoria "other" sisällön selvittäminen!

Yleisesti eri kuvien tekstit yms. tsekkaaminen, korostukset yms.

---

## Global knowledge production and library catalogues

**Idea**: Study early modern knowledge production, all of it (1470-1830)

**How?** Combine library catalogues (Fennica, Kungliga, ESTC, CERL HPB database)

**Important**: cleaning up the data (90% of the effort!) and considering how particular data collections (Finnish National Bibliography, English Short-Title Catalogue etc.) have been formed.

**How to fail**: take worldcat or similar combination of datasets and analyse it as a whole without cleaning it up first.

**Overcoming a national gaze** focus on knowledge production through more complex cross-national networks

---

## Data: bibliographic catalogues 1640-1828

**Fennica** Finnish National bibliography

- 900,000+ books and monographies (printed and electronic; 1488-)
- 70,000+ continuous publications (journals or series; 1771-)
- Series, maps, audiovisual, electronic material
- `r nrow(subset(df0, catalog == "Fennica" & publication_year >= 1640 & publication_year < 1828))` documents 1640-1828

**Kungliga**  Swedish National bibliography

- `r nrow(subset(df0, catalog == "Kungliga" & publication_year >= 1640 & publication_year < 1828))` documents 1640-1828
- Altogether > 18 million entries

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=3, echo=FALSE}
library(dplyr)
library(ggplot2)
theme_set(theme_bw(20))
df <- df0 %>% filter(publication_year >=1640 & publication_year <=1828) %>% group_by(catalog, publication_decade) %>% tally()
p <- ggplot(df, aes(x = publication_decade, y = n, color = catalog)) + 
       geom_point() + 
       geom_line() + 
       ylab("Documents (n)") + xlab("Publication year") +
       scale_color_manual(values = c("blue", "darkgreen"))
print(p)
```

---

### Methods: Open data analytical ecosystem

**Customized** dedicated R packages for bibliographic analysis combine flexibility with automation

**Automated** scales well with new data and unforeseen changes; a million documents ~ 30 minutes

**Open source** fully transparent and reproducible data analytics; every single detail, from raw data to final visualization https://github.com/rOpenGov/fennica

**Collaboration & Reuse** further use and enrichment

**Research potential** greatly underestimated; world is full of metadata collections

![Data Science Workflow](data-science.png)  
\tiny{(Figure by Hadley Wickham)}

---

### Data processing is a major effort and should not be underestimated

We have built custom tools to automate all analyses from raw data to final summary documents (these slides, for instance).

![](parsing.png) ![](iceberg.png)

---

### Preprocess & Enrich

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(bibliographica, quiet = TRUE)
```

**Clean up messy entries**

```{r, message=FALSE, warning=FALSE,echo=TRUE}
polish_physical_extent("iii-xxiv, 118, [2] p.")$pagecount
polish_place("Sur l'imprimea londres")
```

**Augment data** (geocoordinates, gender, ..)

```{r, message=FALSE, warning=FALSE}
get_country("Porvoo")
```

---

### Analyse & Visualize

**Fast time line analysis**

```{r, fig.width=10, fig.height=7, message=FALSE, warning=FALSE}
titlecount_timeline(subset(df0, catalog == "Fennica"), 
          "publisher", nmin = 500, mode = "n")$plot
```

---

### Practical exercise: Quantitative Study of Finnish Knowledge Production, 1640-1828

1) General trends & political events

2) Publishers in Sweden and Turku

3) Languages & religious texts in Finland

4) University curriculum & Enlightenment in Finland

Visible roots of the diversity of "Finnishness" reflected in knowledge production, 1640-1828.

---

### 1. General trends & political events

---

[General trends & political events]

### Turku as a Swedish University Town among Uppsala, Tartu and Stockholm

TODO Karttakuvassa Greifswald, Lund ei näy. Tätä hyvä saada kehitettyä, tällaisena ei toimi. LL: ei näy sen takia että näytettynä on vuosikymmen 1650 jolloin Lundissa on 0 dokumenttia ja Greifswaldissa 1. Muissa kymmeniä tai satoja. Data näkyy kuten pitääkin.

```{r map1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=5, eval=FALSE}
library(dplyr)
library(ggmap)
library(ggplot2)
library(gisfin)
source("funcs.R")
theme_set(get_theme_map())
mapdata <- filter(df0,
          (catalog == "Kungliga" &
	      publication_place %in% c("Stockholm", "Uppsala",
	                         "Greifswald", "Lund", "Tartu"))|
	  (catalog == "Fennica" &
	      publication_place == "Turku")) %>%
      group_by(publication_decade, publication_place,
	       latitude, longitude) %>%
      summarise(n = n())

mapdata$size <- mapdata$n
mapdata$highlight <- rep("black", nrow(mapdata))
mapdata$highlight[mapdata$publication_place %in% c("Turku")] <- "red"
mapdata$highlight <- factor(mapdata$highlight)

region <- "Europe.north"
mymap <- get_map(location=geobox(region), color = "bw",
                  source="google",  maptype="terrain")

p0 <- ggmap(mymap) + theme(legend.position="none")

pics <- list()
for (i in c(1650, 1710, 1760, 1800)) {

    # Plot map
    p <- p0

    # Pick the investigated period (sliding window)
    # Sum up the years within this sliding window for each element    	    
    dfw <- subset(mapdata, publication_decade == i) %>%
    	     group_by(publication_place, latitude, longitude, highlight) %>%
		       summarize(n = sum(size))

    if (nrow(dfw) > 0) {
      p <- p0 +
      	geom_point(data = dfw,
      	  aes(x = longitude, y = latitude,
	      size = 20*log10(1+n),
	      color = highlight)) +
	      #color = highlight), alpha = 0.8) +	      
	  ggtitle(i) +
	  scale_size(range = c(2, 15)) +
	  theme(title = element_text(size = 30)) +
	  scale_color_manual(values = c("black", "red"))	  
    }

  pics[[as.character(i)]] <- p

}

library(gridExtra)
grid.arrange(pics[[1]], nrow = 1)
```

---

[General trends & political events]

### Publishing activity over time

- 1700-1721 Fennica drop
- 1760-1770s Kungliga peak 
- 1803-1815 Napoleonic wars

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4, out.width="200px"}
library(magrittr)
library(ggplot2)
library(dplyr)
df2 <- df0 %>% group_by(publication_decade, catalog) %>%
               summarise(n = n())
df2$highlight <- rep(FALSE, nrow(df2))
df2$highlight[df2$catalog == "Kungliga" &
	      df2$publication_decade %in% c(1770, 1810)] <- TRUE
df2$highlight[df2$catalog == "Fennica" &
	      df2$publication_decade %in% c(1710, 1810)] <- TRUE
v <- seq(1500, 1800, 50)
theme_set(theme_bw(20))
p <- ggplot(df2, aes(x = publication_decade, y = n)) +
     geom_line(aes(group = catalog), color = "black") +
     #geom_smooth(aes(fill = catalog)) + 
     geom_point(aes(shape = catalog, color = highlight, size = highlight)) +
     ylab("Title count (n)") + xlab("Publication year") +
     scale_color_manual(values = c("black", "black")) +
     #scale_color_manual(values = c("black")) +
       theme(legend.position = "none") +
     scale_x_continuous(breaks = v, labels = v)
     # ggtitle("Overall publishing activity")
print(p)
```

---

[General trends & political events]

### Comparing publishing activity in Stockholm and Turku

```{r publishingactivitycomparisons, echo=FALSE, message=FALSE, cache=TRUE, fig.width=8, fig.height=5, fig.show="hold", out.width="150px"}
v <- seq(1500, 1800, 100)
selected.places = c("Turku", "Stockholm")
df2 <- df0 %>% group_by(publication_decade, publication_place, catalog) %>%
               summarise(n = n()) %>%
	       filter(publication_place %in% selected.places) 
df2$catalog = factor(df2$catalog, levels = rev(c("Fennica", "Kungliga")))
df2$publication_place = droplevels(factor(df2$publication_place))
df3 = as.data.frame(spread(df2, publication_decade, n, fill = 0))
df2 = melt(df3)
colnames(df2) = c("publication_place", "catalog", "publication_decade", "n")
df2$publication_decade = as.numeric(as.character(df2$publication_decade))
df2$n = as.numeric(as.character(df2$n))
for (catal in unique(df2$catalog)) {
  p <- ggplot(subset(df2, catalog == catal), aes(x = publication_decade, y = n)) +
     geom_line(aes(linetype = publication_place)) +
     geom_point(aes(shape = publication_place), size = 3) +
     ylab("Title count (n)") + xlab("Publication year") +  
     ggtitle(catal) +
     scale_x_continuous(breaks = v, labels = v) + 
     theme(axis.text.x = element_text(angle = 0))
  print(p)

}
```

---

[General trends & political events]

### Explaining the 1760-1770 publishing peak in Sweden

Pamphleteering plays a crucial role during abolished censorship in Sweden.

```{r pamflets1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4, out.width="280px", fig.show="hold"}
df <- df0 %>%
        filter(catalog == "Kungliga")

# Add indicator of the publication period
df$period <- cut(df$publication_year,
           breaks = c(1757, 1766, 1775, 1783),
		 include.lowest = TRUE,
		 right = FALSE
		 )

# Remove entries outside the investigated publication periods
df <- df %>% filter(!is.na(period)) %>%
	     group_by(period) %>%
	     summarise(n = n(),
		  paper = sum(paper.consumption.km2, na.rm = TRUE)) %>%
        mutate(paper.per.title = paper/n)

p <- ggplot(df,
       aes(x = period, y = paper.per.title)) + 
       geom_bar(stat = "identity") +
       scale_x_discrete(labels = c("1757-1765", "1766-1774", "1775-1783")) +
       ylab("Paper per title") + xlab("") 
print(p)
```

---

[General trends & political events]

### Stockholm & democratic cycles

An example how historical events, namely parliamentary assemblies in the Swedish case, have an impact on local knowledge production.

TODO: Vaihdettiin tähän Stockholm, riittää julkaisumäärät.

```{r riksdar2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
  myplace <- "Stockholm" # Norrköping / Stockholm
  minyear <- 1700
  maxyear <- 1800
  catalog <- "Kungliga"

  df <- df0
  df = subset(df, catalog == catal)
  df <- subset(df, publication_year >= minyear & publication_year <= maxyear)

  # Use timeinterval year intervals
  # Remove special chars
  if (length(unique(df$publication_place[grep("^G.vle$", df$publication_place)])) == 1) {
    df$publication_place[grep("^G.vle$", df$publication_place)] <- "Gavle"
  }
  if (length(unique(df$publication_place[grep("^Norrk.ping$", df$publication_place)])) == 1) {
    df$publication_place[grep("^Norrk.ping$", df$publication_place)] <- "Norrkoping"
  }
  df <- df %>% filter(publication_place == myplace)

  timeinterval <- 1
  df$timeunit <- round(df$publication_year/timeinterval)*timeinterval

  df$unity = rep(1, nrow(df))
  publications <- tapply(df$unity, list(df$timeunit), sum)

  publications[is.na(publications)] <- 0 # Set NAs to 0
  publications <- publications/timeinterval # Instead of decadal sum, use average annual output 
library(reshape2)  
dfm <- melt(publications) 
  names(dfm) <- c("Time", "Documents")
  dfm <- transform(dfm, date = as.numeric(as.character(Time)))
  ymin = min(0.1)
  ymax = max(60)

  rect_left <- c(min(na.omit(dfm$date)),
               1719-.5, 1719+.5, # Stockholm 20 januari 1719 1 juni 1719
               1734-.5, 1734+.5, # Stockholm 14 maj 1734 14 december 1734
               1765-.5, 1766+.5, # Stockholm 21 februari 1765 21 oktober 1766
               1769-.5, 1770+.5, # Norrkoping & Stockholm 22 april 1769 5 februari 1770
               1771-.5, 1772+.5, # Stockholm 19 juni 1771 12 september 1772
               1792-.5, 1792+.5, # Gavle 26 januari 1792 24 februari 1792	       
               max(na.omit(dfm$date)))
  rectangles <- data.frame(
    xmin = rect_left[-length(rect_left)],
    xmax = rect_left[-1],
    ymin = ymin,
    ymax = ymax
    )
  rectangles$shade <- rep(c("White", "Highlight"), length = nrow(rectangles))

 riksplace = c("Stockholm", "Stockholm", "Stockholm", "Norrkoping", "Stockholm", "Gavle")
 cols = c("gray", "lightblue", "yellow", "white")
 rectangles$shade[rectangles$shade == "Highlight"] = riksplace
 rectangles$shade = factor(rectangles$shade)

  # Draw Figure
library(ggplot2)
  theme_set(theme_bw(20))
  p <- ggplot()
  p <- p + geom_rect(data = rectangles, 
	   aes(xmin=xmin, xmax=xmax,
	       ymin=ymin, ymax=ymax,
	       fill=shade), alpha=0.8) + 
         scale_fill_manual(values = cols) # + guides(fill = "none") 
  p <- p + geom_line(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + geom_point(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + ggtitle("Publishing activity")
  p <- p + ylab("Documents / Year")
  p <- p + xlab("Year")
  p <- p + ggtitle(paste(myplace, " (", catal, ")", sep = ""))
  print(p)
```

---

### 2. Publishers in Sweden and Turku

---

[Publishers in Sweden and Turku]

### Publishers in Swedish University towns (and Finland)

```{r publishers2-finland, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
df <- df0
df <- df %>%
        filter(catalog == "Fennica") %>%
	filter(country == "Finland" | publication_place == "Vyborg") %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

# Top publication places
library(bibliographica)

top <- c("Turku", "Vaasa", "Vyborg")
df <- df %>%	
        filter(publication_place %in% top) %>%	
        select(publication_decade, publication_place, publisher)

npub <- unique(df) %>% group_by(publication_decade, publication_place) %>% tally()
npub$publication_place <- factor(npub$publication_place, levels = top)
npub.fennica <- npub

df <- df0
catalogue <- "Kungliga"
df <- df %>%
        filter(catalog == catalogue) %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

# Selected publication places
top <- c("Stockholm", "Lund", "Uppsala", "Greifswald", "Tartu")
df <- df %>%	
        filter(publication_place %in% top) %>%	
        select(publication_decade, publication_place, publisher)

npub <- unique(df) %>% group_by(publication_decade, publication_place) %>% tally()
npub$publication_place <- factor(npub$publication_place, levels = top)
npub.kungliga <- npub

npub <- bind_rows(npub.fennica, npub.kungliga)

# Order by total publishing activity
npub$publication_place <- factor(npub$publication_place, levels = (npub %>% group_by(publication_place) %>% summarise(total = sum(n)) %>% arrange(desc(total)))$publication_place)
npub <- npub %>% arrange(publication_place)

theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack",
       		    aes(fill = publication_place)) + 
       xlab("Publication year") +
       ylab("Unique publishers (n)") +        
       guides(fill = guide_legend(reverse = TRUE, title = "")) #+ 
       #ggtitle(paste("Unique publishers in selected publication places (", catalogue, ")"))

print(p)
```

---

[Publishers in Sweden and Turku]

### Turku in Sweden: proportions for title count and paper in Swedish University towns

```{r comparisons, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=18, fig.height=10}
pics <- list()
for (catalogue in na.omit(unique(df0$catalog))) {

df <- df0 %>% filter(catalog == catalogue)%>%
        filter(publication_year >= 1640 & publication_year <= 1828)
places <- c("Stockholm", "Uppsala", "Lund", "Greifswald", "Turku")
df$publication_place[!df$publication_place %in% places] <- "Other"
df <- subset(df, !is.na(df$publication_place))
#df$publication_place <- factor(df$publication_place, levels = c(places, "Other"))
library(reshape2)
library(magrittr)
library(tidyr)

df <- df %>% group_by(publication_decade, publication_place) %>%
             summarise(n = n(),
       	       paper = sum(paper.consumption.km2, na.rm = TRUE),
		       publishers = length(na.omit(unique(publisher)))
		       )

# Calculate percentages
for (varname in c("n", "paper", "publishers")) {

  df$varname <- df[[varname]]
  dff <- df %>% select(publication_decade, publication_place, varname)
  dff <- spread(dff, publication_place, varname, fill = 0)
  dff[, -1] <- 100 * t(apply(dff[, -1], 1, function (x) {x/sum(x)}))
  dff <- melt(dff, "publication_decade");
  colnames(dff) <- c("publication_decade", "publication_place", "f")

  theme_set(theme_bw(20))
  p <- ggplot(dff, aes(x = publication_decade, y = f)) +
     geom_bar(position = "stack", stat = "identity",
       aes(fill = publication_place)) + 
     xlab("Publication year") +
     scale_fill_brewer(palette="Paired")
     if (varname == "n") {
       p <- p + ggtitle(paste("Title count (", catalogue, ")", sep = "")) + 
                ylab("Title count frequency (%)") 
     }
     if (varname == "paper") {
       p <- p + ggtitle(paste("Paper (", catalogue, ")", sep = "")) +
                ylab("Paper (%)")        
     }
     if (varname == "publishers") {
       p <- p + ggtitle(paste("Publishers (", catalogue, ")", sep = "")) +
                       ylab("Publishers (%)")
     }     
  p <- p + guides(fill = guide_legend(reverse = TRUE, title = ""))

  pics[[paste(catalogue, varname, sep = "-")]] <- p

}
}
library(gridExtra)
grid.arrange(pics[[1]], pics[[2]], pics[[4]], pics[[5]], nrow = 2)
rm(pics)
```

---

[Publishers in Sweden and Turku]

### Top publishers in Turku/Fennica (title count)

```{r publishers4-fennica, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
place <- "Turku"
catalogue = "Fennica"
df = df0

df <- df %>%
        filter(catalog == catalogue & publication_place == place)

df <- df %>%	
  filter(publication_year >= 1640 & publication_year <= 1828)%>%
        select(publication_decade, publisher, paper.consumption.km2)

# Group small publishers
# Top publishers by title count
ntop <- 10
top <- names(top(df, "publisher", ntop))
df$publisher[!df$publisher %in% top] <- "Other"
df$publisher <- factor(df$publisher, c(top, "Other"))

# Title count per decade & publisher	
npub <- df %>% group_by(publication_decade, publisher) %>% tally()

# TITLE COUNT
theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack", aes(fill = publisher)) + 
       xlab("Publication year") +
       ylab("Title count (n)") #+        
       #ggtitle(paste("Title count per publisher (", place, "/", catalogue, ")", sep = ""))
print(p)
```

---

### 3. Languages & religious texts in Finland

---

[Languages & religious texts in Finland]

### Language comparison between catalogues (percentage)

```{r languageperc, echo=FALSE, message=FALSE, warning=FALSE, fig.width=15, fig.height=10}
pics <- list()
for (catalogue in c("Fennica", "Kungliga")) {

  if (catalogue == "Fennica") {
    df <- df.full # full time span
  } else if (catalogue == "Kungliga") {
    df <- df.full # full time span
  }
  langs <- c("Finnish", "Swedish", "Latin", "German", "Russian", "French", "Other")
  df <- df %>% filter(catalog == catalogue)%>%
        filter(publication_year >= 1600 & publication_year <= 1900)

library(dplyr)
library(tidyr)
library(magrittr)
library(reshape2)

lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0

dfl <- NULL
for (lan in lang) {

  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  if (nrow(dflsub)>0) {
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}

df <- dfl %>% group_by(publication_decade, language) %>%
     	     summarise(n = n())

# Calculate percentages
dff <- spread(df, language, n, fill = 0)
dff[, -1] <- 100 * t(apply(dff[, -1], 1, function (x) {x/sum(x)}))
dff <- melt(dff, "publication_decade");
colnames(dff) <- c("publication_decade", "language", "f")

theme_set(theme_bw(20))
p <- ggplot(dff, aes(x = publication_decade, y = f)) +
     geom_bar(position = "stack", stat = "identity", aes(fill = language)) + 
     xlab("Publication year") +
     ylab("Title count frequency (%)") +
     #ggtitle(paste("Languages (", catalogue, ")", sep = "")) +
     ggtitle(catalogue) +
     guides(fill = guide_legend(reverse = TRUE, title = "")) +
     scale_fill_brewer(palette="Paired")
pics[[catalogue]] <- p
}
library(gridExtra)
grid.arrange(p[[1]], p[[2]], nrow = 2)
```

---

[Languages & religious texts in Finland]

### Different languages in Turku, Stockholm, Lund and Uppsala

```{r language-perc2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=15}
pics <- list()
theme_set(theme_bw(20))
myplaces <- c("Turku", "Greifswald", "Stockholm", "Uppsala", "Lund")

for (catalogue in c("Fennica", "Kungliga")) {

  df <- df0 %>% filter(catalog == catalogue) %>%
        filter(publication_year >= 1600 & publication_year <= 1828)

  lang <- paste("language.", langs, sep = "")
  otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
  df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0

  dfl <- NULL
  for (lan in lang) {

    # Classify a document to the specifed language
    # If document is assigned with languages, each case is considered
    # so one doc may have multiple entries corresponding to different languages
    # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
    # Combine data frames for specified languages
    dflsub <- filter(df, df[[lan]])
    if (nrow(dflsub) > 0) {
      dflsub$language <- gsub("language.", "", lan)
      dfl <- bind_rows(dfl, dflsub)
    }
  }

  df <- dfl %>% group_by(publication_decade, publication_place, language) %>%
     	     summarise(n = n())

  # Calculate percentages
  dff <- spread(df, language, n, fill = 0)
  dff[, -c(1,2)] <- 100 * t(apply(dff[, -c(1,2)], 1, function (x) {x/sum(x)}))

  dff <- melt(dff, id = c("publication_place", "publication_decade"))
  colnames(dff) <- c("publication_place", "publication_decade", "language", "f")

  for (myplace in myplaces) {
    p <- ggplot(subset(dff, publication_place == myplace), aes(x = publication_decade, y = f)) +
      geom_bar(position = "stack", stat = "identity", aes(fill = language)) + 
      xlab("Publication year") +
      ylab("Title count frequency (%)") +
      #ggtitle(paste("Languages (", myplace, "/", catalogue, ")", sep = "")) +
      ggtitle(paste(myplace, "/", catalogue)) +
      scale_fill_brewer(palette="Paired")
    pics[[paste(catalogue, myplace, sep = "-")]] <- p
  }
}

grid.arrange(pics[[1]],
             pics[[8]], pics[[9]], pics[[10]], 
             nrow = 2)

rm(pics)
```

---

[Languages & religious texts in Finland]

### Topic richness per language (Finnish publications)

```{r topics232, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
catalogue <- "Fennica"
df <- filter(df0, catalog == "Fennica") %>%
  filter(publication_year >= 1640 & publication_year <= 1828) %>%
  select(publication_year, language, subject_topic)
df <- unique(df)
df <- df %>% group_by(publication_year, language) %>%
             summarise(n = length(unique(unlist(strsplit(subject_topic, ";")))))

df$language[!df$language %in% langs] <- "Other"
df$language <- factor(df$language, levels = langs)

p <- ggplot(df, aes(x = publication_year, y = n, color = language)) +
       geom_line() + geom_point() +
       ggtitle("Number of subject topics")

print(p)
```

---

[Languages & religious texts in Finland]

### Devotional literature (catechisms, hymns, prayer collections, etc.)

- Form of literature denoting the advancement of reading in Finland in the nineteenth century. It has been unclear when does it show up as a relevant category.

```{r topics12122, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/topics-2.png
sel <- c("virret","arkkiveisut","hartauskirjat","katekismukset","rukouspäivät","saarnat","aapiset","rukoukset","rukous","hengelliset laulut","hartauspuheet","virsikirjat")
catalogue <- "Fennica"

  # Selected catalogue with selected years
  df <- filter(df0, catalog == catalogue &
     		    publication_year >= 1640 &
		    publication_year <= 1828)

  # Selected topics
  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
  df <- df %>% filter(catalog == catalogue & hit)

  lang <- paste("language.", langs, sep = "")
  otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
  df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
  dfl <- NULL
  for (lan in lang) {
    # Classify a document to the specifed language
    # If document is assigned with languages, each case is considered
    # so one doc may have multiple entries corresponding to different languages
    # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
    # Combine data frames for specified languages
    dflsub <- filter(df, df[[lan]])
    if (nrow(dflsub) > 0) {
      dflsub$language <- gsub("language.", "", lan)
      dfl <- bind_rows(dfl, dflsub)
    }
  }
  df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

  # PAPER CONSUMPTION
  theme_set(theme_bw(20))
  p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Languages (", catalogue, ")", sep = "")) 
  print(p)
```

---

[Languages & religious texts in Finland]

### Temporary sermons in Latin and Swedish

- play an important role in early modern knowledge production. It has not been analysed in what language they were in Finland (and Sweden).

```{r topics882, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
  sel <- c("hautajaiset", "häät", "juhlamenot")
  catalogue <- "Fennica"
  df <- df0
  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)

  # Selected catalogue with selected topics
  df <- df %>% filter(catalog == catalogue & hit)
  lang <- paste("language.", langs, sep = "")
  otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
  df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0

  dfl <- NULL
  for (lan in lang) {
    # Classify a document to the specifed language
    # If document is assigned with languages, each case is considered
    # so one doc may have multiple entries corresponding to different languages
    # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
    # Combine data frames for specified languages
    dflsub <- filter(df, df[[lan]])
    if (nrow(dflsub) > 0) {
      dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}
dfl$language <- factor(dfl$language, levels = langs)

  df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

  # PAPER CONSUMPTION
  p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
  print(p)
```

---

### 4. University curriculum & Enlightenment in Finland

---

[University curriculum & Enlightenment in Finland]

### Dissertation languages in Turku 1640-1828

```{r disslang, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
catalogue <- "Fennica"
langs <- c("Finnish", "Swedish", "Latin", "German",
           "Russian", "French", "Hebrew", "Other")

df <- df0 %>%
        filter(catalog == catalogue) %>%
        filter(publication_year >= 1640 & publication_year <= 1828) %>%
	filter(dissertation)

lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0

dfl <- NULL
for (lan in lang) {

  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  if (nrow(dflsub)>0) {
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}

df <- dfl %>% group_by(publication_decade, language) %>%
     	     summarise(n = n())

# Calculate percentages
dff <- spread(df, language, n, fill = 0)
dff[, -1] <- 100 * t(apply(dff[, -1], 1, function (x) {x/sum(x)}))
dff <- melt(dff, "publication_decade");
colnames(dff) <- c("publication_decade", "language", "f")

# Order by language counts
dff$language <- factor(dff$language,
   levels = (df %>% group_by(language) %>%
                    summarise(total = sum(n)) %>%
		    arrange(total))$language);
dff <- dff %>% arrange(desc(language))

theme_set(theme_bw(20))
p <- ggplot(dff, aes(x = publication_decade, y = f)) +
     geom_bar(position = "stack", stat = "identity", aes(fill = language)) + 
     xlab("Publication year") +
     ylab("Title count frequency (%)") +
     ggtitle(paste("Languages (", catalogue, ")", sep = "")) +
     scale_fill_brewer(palette="Paired")
print(p)
```

---

[University curriculum & Enlightenment in Finland]

### Printing of philosophy in different languages in Turku

```{r topics-103, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel = c("filosofia","luonnonfilosofia","metafysiikka","logiikka")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
                 paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_bar(aes(fill = language), position = "stack", stat = "identity") +                     
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Paper (", paste(sel, collapse = ";"), ")", sep = ""))
print(p)

```

---

[University curriculum & Enlightenment in Finland]

### Printing of history in different languages in Turku

```{r topics-102, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel = c("oppihistoria","antiikki","historia")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_bar(aes(fill = language), position = "stack", stat = "identity") +             
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Paper (", paste(sel, collapse = ";"), ")", sep = ""))
print(p)

```

---

[University curriculum & Enlightenment in Finland]

### Printing of economy and social science in different languages in Turku

```{r topics-105, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel =   c("psykologia","yhteiskuntafilosofia","valtiofilosofia","talous","raha","moraali","velvollisuudet","kasvatus","maanviljely","maatalous","kalastus")

# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_bar(aes(fill = language), position = "stack", stat = "identity") +              
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Paper (", paste(sel, collapse = ";"), ")", sep = ""))
print(p)

```

---

[University curriculum & Enlightenment in Finland]

### Rise of the octavo sized book as vehicle of Enlightenment

Paper consumption for different document formats over time.

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=8, fig.show="hold", out.width="150px"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper.consumption.km2 = sum(paper.consumption.km2, na.rm = TRUE), n = n()) 
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  df2$highlight <- rep("gray", nrow(df2))
  df2$highlight[df2$gatherings == "8vo"] <- "darkgreen"

  p <- ggplot(df2, aes(y = paper.consumption.km2,
                       x = publication_decade,
  	       shape = gatherings,
		       linetype = gatherings))		       
  p <- p + geom_point(size = 4)
  p <- p + geom_smooth(method = "loess", size = 1,
         aes(color = highlight, fill = highlight))
  p <- p + scale_color_manual(values = c("darkgreen", "darkgray"))
  p <- p + scale_fill_manual(values = c("darkgreen", "darkgray"))  	 
  p <- p + ggtitle("Paper consumption in time by gatherings")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption (km2)")
  p <- p + guides(linetype = guide_legend(keywidth = 5),
       	          shape = guide_legend(keywidth = 5))
  p <- p + ggtitle(paste("Paper consumption\n(", catal, ")", sep = ""))
  print(p)
}
```

---

[University curriculum & Enlightenment in Finland]

### Enlightenment literature read in Finland

```{r authors22, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8, out.width="160px", fig.show="hold"}
theme_set(theme_bw(20))
for (catalogue in c("Fennica", "Kungliga")) {
  df <- filter(df0, catalog == catalogue) %>%
        filter(publication_year >= 1640 & publication_year <= 1828)
  top.authors <- names(bibliographica::top(df, field = "author", n = 10))
  dfs <- df %>% filter(author %in% top.authors) %>%
         group_by(author, publication_decade) %>%
         tally() %>%
         arrange(publication_decade)
  p <- ggplot(dfs, aes(x = publication_decade, y = n, fill = author)) +
       geom_bar(stat = "identity", position = "stack", color = "black") +
       xlab("Publication Decade") +
       ylab("Title Count") +
       scale_fill_grey() +
       guides(fill = guide_legend("Author")) +
       ggtitle(paste("Title count top authors\n", catalogue)) +
       xlim(c(1640, 1828))
  print(p)
}
```

---

### Thanks !

R master (plots & maps):
- Leo Lahti

Institutional support:

- Academy of Finland
- University of Helsinki
- Digitalia / The Regional Council of South Savo

Data providers:

- National Library of Finland
- National Library of Sweden

Source code for these slides:
- https://github.com/rOpenGov/fennica/blob/master/inst/examples/20160715-slides-Krakow.Rmd

---

### Duplicate publications in Fennica/Kungliga

```{r duplicates, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=5, fig.width=10}
df <- df.combined.preprocessed %>%
        filter(publication_year >= 1640 & publication_year <=1828) %>%
        group_by(catalog, publication_decade) %>%
	summarise(duplicates = 100 * mean(remove))

p <- ggplot(df, aes(x = publication_decade, y = duplicates, fill = catalog)) +
       geom_bar(position = "dodge", stat = "identity") +
       ylab("Duplicates (%)") +
       scale_fill_manual(values = c("blue", "red")) 
       #ggtitle(paste("Certain duplicates Fennica-Kungliga", paste(range(na.omit(df$publication_decade)), collapse = "-")))

print(p)
```