### 2016-07-12 inst/examples/slides.Rmd
ok

---
title: "Untitled"
output: ioslides_presentation
---

### Data processing is a major effort and should not be underestimated

We have built custom tools to automate all analyses from raw data to final summary documents (these slides, for instance).

\begin{columns}
\begin{column}{0.5\textwidth}
     \includegraphics[width=1\textwidth]{parsing.png}
\end{column}
\begin{column}{0.5\textwidth}  %%<--- here
    \begin{center}
     \includegraphics[width=1\textwidth]{iceberg.png}
     \end{center}
\end{column}
\end{columns}

---

![](ecosystemflowchart.png)