### 2016-07-12 inst/examples/slides.Rmd
sync

---
title: "Untitled"
output: ioslides_presentation
---

### Data processing is a major effort and should not be underestimated

We have built custom tools to automate all analyses from raw data to final summary documents (these slides, for instance).

\begin{columns}
\begin{column}{0.5\textwidth}
     \includegraphics[width=1\textwidth]{parsing.png}
\end{column}
\begin{column}{0.5\textwidth}  %%<--- here
    \begin{center}
     \includegraphics[width=1\textwidth]{iceberg.png}
     \end{center}
\end{column}
\end{columns}

---

![](ecosystemflowchart.png)

---

### Data: bibliographic catalogues 1640-1828

**Fennica** Finnish National bibliography

- 900,000+ books and monographies (printed and electronic; 1488-)
- 70,000+ continuous publications (journals or series; 1771-)
- Series, maps, audiovisual, electronic material
- `r nrow(subset(df0, catalog == "Fennica" & publication_year >= 1640 & publication_year < 1828))` documents 1640-1828

**Kungliga**  Swedish National bibliography

- `r nrow(subset(df0, catalog == "Kungliga" & publication_year >= 1640 & publication_year < 1828))` documents 1640-1828
- Altogether > 18 million entries

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=3, echo=FALSE}
library(dplyr)
library(ggplot2)
theme_set(theme_bw(20))
df <- df0 %>% filter(publication_year >=1640 & publication_year <=1828) %>% group_by(catalog, publication_decade) %>% tally()
p <- ggplot(df, aes(x = publication_decade, y = n, color = catalog)) + 
       geom_point() + 
       geom_line() + 
       ylab("Documents (n)") + xlab("Publication year") +
       scale_color_manual(values = c("blue", "darkgreen"))
print(p)
```

---

## Analyse & Visualize

**Fast time line analysis**

```{r, fig.width=15, fig.height=5, message=FALSE, warning=FALSE}
library(bibliographica)
titlecount_timeline(subset(df0, catalog == "Fennica"), 
          "publisher", nmin = 500, mode = "n")$plot
```

---