### 2017-01-18 inst/examples/Gaudeamus.Rmd

---
title: "Gaudeamus Figures"
author: "Leo Lahti and Mikko Tolonen"
date: "`r Sys.Date()`"
output: markdown_document
---

```{r, echo=FALSE, message=FALSE}
library(dplyr)
knitr::opts_chunk$set(fig.path = "figure_Gaudeamus/", dev="CairoPNG")
dfo <- readRDS("df.Rds")
theme_set(theme_bw(40))
```

```{r Vaasa, echo=FALSE, fig.width=20, fig.height=10}
df <- dfo
place <- "Vaasa"
miny <- 1764
maxy <- 1917
df <- df %>% filter(!is.na(publication_year)) %>%
             filter(publication_year >= miny & publication_year <= maxy)
tmp1 <- df %>% filter(publication_place == place) %>%
     	       group_by(publication_decade, publication_year) %>%
	       tally() %>%
	       arrange(publication_year)

#tmp1$decade <- as.numeric(paste0(substr(tmp1$publication_year,1,3),"0"))
tmp2 <- tmp1 %>% group_by(publication_decade) %>% summarise(m = sum(n))

years <- c(1764,1809,1863,1917)
p <- ggplot(tmp1, aes(x = publication_year, y = n)) +
#       geom_point() +
       geom_bar(aes(x = publication_decade + 5, y = m), stat = "identity", fill = "gray", alpha = 0.9, color = "black", data = tmp2) +
       ggtitle(paste("Julkaisutoiminta Vaasassa ", miny, "-", maxy, sep = "")) +
       ylab("Dokumenttia (n)") +
       scale_x_continuous(labels = years, breaks = years) +
       xlab("Julkaisuvuosi")

print(p)
```

```{r Turku, echo=FALSE, fig.width=20, fig.height=10}
df <- dfo
place <- "Turku"
miny <- 1640
maxy <- 1917
df <- df %>% filter(!is.na(publication_year)) %>%
             filter(publication_year >= miny & publication_year <= maxy)
tmp1 <- df %>% filter(publication_place == place) %>%
     	       group_by(publication_decade, publication_year) %>%
	       tally() %>%
	       arrange(publication_year)

#tmp1$decade <- as.numeric(paste0(substr(tmp1$publication_year,1,3),"0"))
tmp2 <- tmp1 %>% group_by(publication_decade) %>% summarise(m = sum(n))

years <- c(1640, 1764,1809,1863,1917)
p <- ggplot(tmp1, aes(x = publication_year, y = n)) +
#       geom_point() +
       geom_bar(aes(x = publication_decade + 5, y = m), stat = "identity", fill = "gray", alpha = 0.9, color = "black", data = tmp2) +
       ggtitle(paste("Julkaisutoiminta Turussa ", miny, "-", maxy, sep = "")) + 
       ylab("Dokumenttia (n)") +
       scale_x_continuous(labels = years, breaks = years) +
       xlab("Julkaisuvuosi")

print(p)
```

```{r Helsinki, echo=FALSE, fig.width=20, fig.height=10}
df <- dfo
place <- "Helsinki"
miny <- 1746
maxy <- 1917
df <- df %>% filter(!is.na(publication_year)) %>%
             filter(publication_year >= miny & publication_year <= maxy)
tmp1 <- df %>% filter(publication_place == place) %>%
     	       group_by(publication_decade, publication_year) %>%
	       tally() %>%
	       arrange(publication_year)

#tmp1$decade <- as.numeric(paste0(substr(tmp1$publication_year,1,3),"0"))
tmp2 <- tmp1 %>% group_by(publication_decade) %>% summarise(m = sum(n))

years <- c(1764,1809,1863,1917)
p <- ggplot(tmp1, aes(x = publication_year, y = n)) +
#       geom_point() +
       geom_bar(aes(x = publication_decade + 5, y = m), stat = "identity", fill = "gray", alpha = 0.9, color = "black", data = tmp2) +
       ggtitle(paste("Julkaisutoiminta Helsingissä ", miny, "-", maxy, sep = "")) +

       ylab("Dokumenttia (n)") +
       scale_x_continuous(labels = years, breaks = years) +
       xlab("Julkaisuvuosi")

print(p)
```

```{r Suomi, echo=FALSE, fig.width=20, fig.height=10}
df <- dfo
place <- "Finland"
miny <- 1640
maxy <- 1917
df <- df %>% filter(!is.na(publication_year)) %>%
             filter(publication_year >= miny & publication_year <= maxy)
tmp1 <- df %>% filter(country == place) %>%
     	       group_by(publication_decade, publication_year) %>%
	       tally() %>%
	       arrange(publication_year)

#tmp1$decade <- as.numeric(paste0(substr(tmp1$publication_year,1,3),"0"))
tmp2 <- tmp1 %>% group_by(publication_decade) %>% summarise(m = sum(n))

years <- c(1640, 1764,1809,1863,1917)
p <- ggplot(tmp1, aes(x = publication_year, y = n)) +
#       geom_point() +
       geom_bar(aes(x = publication_decade + 5, y = m), stat = "identity", fill = "gray", alpha = 0.9, color = "black", data = tmp2) +
       ggtitle(paste("Julkaisutoiminta Suomessa ", miny, "-", maxy, sep = "")) +               
       ylab("Dokumenttia (n)") +
       scale_x_continuous(labels = years, breaks = years) +
       xlab("Julkaisuvuosi")

print(p)
```

```{r Kaikki, echo=FALSE, fig.width=20, fig.height=10}
df <- dfo
countries <- "Finland"
places <- c("Helsinki", "Turku", "Vaasa")
miny <- 1640
maxy <- 1917
df <- df %>% filter(!is.na(publication_year)) %>%
             filter(publication_year >= miny & publication_year <= maxy)

tmp1 <- df %>% filter(publication_place %in% places) %>%
     	       group_by(publication_decade, publication_place) %>%
	       tally()

tmp2 <- df %>% filter(country %in% countries) %>%
     	       group_by(publication_decade, country) %>%
	       tally() %>%
	       rename(publication_place = country)

d <- bind_rows(tmp1, tmp2)
d$publication_place <- factor(d$publication_place)

years <- c(1640, 1764,1809,1863,1917)
p <- ggplot(d, aes(x = publication_decade + 5, y = n, fill = publication_place, color = publication_place)) +
      # geom_bar(stat = "identity", alpha = 0.9, color = "black", position = "dodge") +
       geom_smooth() +       
       ggtitle("Julkaisutoiminta 1640-1917") +
       ylab("Dokumenttia (n)") +
       scale_x_continuous(labels = years, breaks = years) +
       xlab("Julkaisuvuosi") +
       scale_y_log10()

print(p)
```