### 2017-01-26 inst/examples/20170125-Manuscript-FK.Rmd
Fennica analyses mod

---
title: "Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828"
author: "Mikko Tolonen, Jani Marjanen, Hege Roivainen, Leo Lahti"
date: "`r Sys.Date()`"
output: 
  beamer_presentation:
    theme: "boxes"
    colortheme: "orchid"
    fonttheme: "professionalfonts"
  pdf_document:
fontsize: 13pt
---

```{r jm-init, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Default time span
min.year <- 1488
max.year <- 1828

library(ggmap)
library(gisfin)
library(stringr)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)
library(devtools)
library(magrittr)
library(ggplot2)
library(dplyr)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.path = "2017-manuscript/")

# Set locale
tmp <- Sys.setlocale(locale="UTF-8")

# Nice theme
theme_set(theme_bw(26))
```

```{r jm-init2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Full combined catalogue (Fen + Kun) with marked duplicates
df.combined.preprocessed <- readRDS("df.combined.20160715-Krakow.Rds")
# Data with duplicates removed and years limited
df0 <- df.combined.preprocessed %>% filter(!remove) %>%
	   			    filter(publication_year >= min.year &
					   publication_year <= max.year)
```

### Paper consumption per title

Paper consumption per title in 1757–1765, 1766–1774, and 1775–1783 in
Stockholm, Lund and Uppsala in Kungliga.

```{r paper_per_title, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=4}
catalogue <- "Kungliga"
places <- c("Stockholm", "Lund", "Uppsala")
periods <- list(c(1757, 1765), c(1766, 1774), c(1775, 1783))

df <- df0
df$period <- rep(NA, nrow(df))
for (i in 1:length(periods)) {
  pr <- periods[[i]]
  df$period[df$publication_year >= min(pr) & df$publication_year <= max(pr)] <- paste(min(pr), max(pr), sep = "-")
}
df$period <- factor(df$period)

# Selected catalogue with selected years
df <- filter(df, catalog == catalogue & publication_place %in% places & !is.na(period) & !is.na(paper))

df <- df %>% group_by(period, publication_place) %>%
             summarise(n = n(), paper = sum(paper, na.rm = TRUE)) %>%
	     mutate(paper_per_title = paper/n)

theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_place, y = paper_per_title)) +
     geom_bar(stat = "identity", position = "dodge", aes(fill = period), color = "black") +
     ylab("Paper per title (sheets)") +
     xlab("Publication period") +
     #scale_color_manual(values = c("blue", "darkgreen")) + 
     # 	 scale_size_discrete(range = c(2, 8)) +
     guides(fill = guide_legend(reverse = TRUE, title = "Place"), color = "none") 
print(p)

kable(df)
```

### Book production and Riksdag assemblies

According to Kungliga in the the long eighteenth century.

```{r riksdar, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
minyear <- 1690
maxyear <- 1828 # or 1810
catal <- "Kungliga"

library(magrittr)
library(ggplot2)
library(dplyr)
library(ggplot2)
library(reshape2)

  df <- df0
  df <- subset(df, catalog == catal)
  df <- subset(df, publication_year >= minyear & publication_year <= maxyear)

  timeinterval <- 1
  df$timeunit <- round(df$publication_year/timeinterval)*timeinterval

  df$unity = rep(1, nrow(df))
  publications <- tapply(df$unity, list(df$timeunit), sum)

  publications[is.na(publications)] <- 0 # Set NAs to 0
  publications <- publications/timeinterval # Instead of decadal sum, use average annual output 
  dfm <- melt(publications) 
  names(dfm) <- c("Time", "Documents")
  dfm <- transform(dfm, date = as.numeric(as.character(Time)))
  ymin = min(dfm$Documents)
  ymax = max(dfm$Documents)

  li <- list(1719, # Stockholm 20 januari 1719 1 juni 1719
             1720,
	     1723,
	     c(1726, 1727),
	     1731,
	     1734, # Stockholm 14 maj 1734 14 december 1734
	     c(1738, 1739),
	     c(1740, 1741),
	     1742,
	     c(1746, 1747),
	     c(1751, 1752),
	     c(1755, 1756),
	     c(1760, 1762),
	     c(1765, 1766), # Stockholm 21 februari 1765 21 oktober 1766
	     c(1769, 1770), # Norrköping & Stockholm 22 april 1769 5 februari 1770
	     c(1771, 1772), # Stockholm 19 juni 1771 12 september 1772
	     c(1778, 1779),
	     1786,
	     1789,
	     1792, # Gävle 26 januari 1792 24 februari 1792         
	     1800,
	     c(1809, 1810))

  rect_left <- c(min(na.omit(dfm$date)))
  for (item in li) {
    rect_left <- c(rect_left, min(item) - .5, max(item) + .5)
  }
  rect_left <- c(rect_left, max(na.omit(dfm$date)))

  rectangles <- data.frame(
    xmin = rect_left[-length(rect_left)],
    xmax = rect_left[-1],
    ymin = ymin,
    ymax = ymax
    )
  rectangles$shade <- rep(c("Background", "Highlight"), length = nrow(rectangles))

 cols <- c("lightgray", "darkgray")
 bgcol <- "white"
 rectangles$shade[rectangles$shade == "Background"] <- bgcol
 rectangles$shade[rectangles$shade == "Highlight"] <- rep(cols, length.out = sum(rectangles$shade == "Highlight"))
 rectangles$shade = factor(rectangles$shade)

  # Draw Figure
  theme_set(theme_bw(20))
  p <- ggplot()
  p <- p + geom_rect(data = rectangles, 
       aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, fill=shade), alpha=0.8) + 
         scale_fill_manual(values = levels(rectangles$shade)) # + guides(fill = "none") 
  p <- p + geom_line(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + geom_point(data = dfm, aes(x = date, y = Documents), col = "black")
  #p <- p + scale_x_continuous(breaks=seq(min(dfm$date), max(dfm$date), 10))
  p <- p + scale_x_continuous(breaks=scales::pretty_breaks(n = 20))  
  p <- p + ggtitle("Publishing activity")
  p <- p + ylab("Documents / Year")
  p <- p + xlab("Year")
  p <- p + ggtitle(catal)
  p <- p + guides(fill = "none") 
  print(p)
```

### Book production

Book production by year in Turku in Fennica and Kungliga 1640–1828 as a percentage of all books in the catalogues.

```{r bookprod, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
minyear <- 1640
maxyear <- 1828 
df <- df0 %>% filter(!is.na(publication_place)) 
df$myplace <- df$publication_place
df$myplace[!df$myplace == "Turku"] <- "Other"

df <- df %>%  filter(publication_decade >= minyear & publication_decade <= maxyear) %>%
	      group_by(catalog, publication_decade, myplace) %>%
	      summarize(n = n())

pics <- list()
for (cat in unique(df$catalog)) {
  dfs <- subset(df, catalog == cat)
  p <- ggplot(dfs, aes(x = publication_decade, y = n)) +
       geom_bar(aes(fill = myplace), color = "black", stat = "identity", position = "stack") +
       scale_fill_manual(values = c("gray", "black")) +
       ylab("Title count") +
       xlab("Publication decade") +
       guides(fill = guide_legend(title = "Place")) +
       ggtitle(cat) +
       xlim(values = c(min(df$publication_decade), max(df$publication_decade)))
  pics[[cat]] <- p
}

library(gridExtra)
grid.arrange(pics[[1]], pics[[2]], nrow = 2)
```