### 2018-02-15 inst/examples/201705-Gaudeamus.Rmd
Gaudeamus polish

---
title: "Gaudeamus 2017 Figures"
author: "Mikko Tolonen, Leo Lahti"
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    dev: cairo_pdf  
  pdf_document:
    dev: cairo_pdf
    keep_tex: yes
fontsize: 13pt
---

```{r init, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Data files
input.data.file <- "df.combined.Gaudeamus2017.Rds"

# Default time span
library(ggmap)
library(gisfin)
library(stringr)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)
library(devtools)
knitr::opts_chunk$set(echo = FALSE)
# knitr::opts_chunk$set(fig.path = "figure-201705-Gaudeamus/")
# knitr::opts_chunk$set(fig.path = "slides_201606_Krakow", dev="CairoPNG")
# knitr::opts_chunk$set(fig.path = "figure-201705-Gaudeamus/", dev="CairoJPEG")
knitr::opts_chunk$set(fig.path = "figure-201705-Gaudeamus/", dev="CairoPDF")

# Set locale
tmp <- Sys.setlocale(locale="UTF-8")

# Nice theme
theme_set(theme_bw(26))
```

```{r init2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Full combined catalogue (Fen + Kun) with marked duplicates
df.combined.preprocessed <- readRDS(input.data.file)
# Data with duplicates removed and years limited
min.year <- 1488
max.year <- 1828
df0 <- df.combined.preprocessed %>% filter(!remove) %>%
	   			    filter(publication_year >= min.year &
					   publication_year <= max.year)
```

---

### Data: bibliographic catalogues 1640-1828

**Fennica** Finnish National bibliography

- `r nrow(subset(df0, catalog == "Fennica" & publication_year >= 1640 & publication_year <= 1828))` documents 1640-1828

**Kungliga**  Swedish National bibliography

- `r nrow(subset(df0, catalog == "Kungliga" & publication_year >= 1640 & publication_year <= 1828))` documents 1640-1828

```{r Gaudeamus-lineplot, message=FALSE, warning=FALSE, fig.width=10, fig.height=3, echo=FALSE}
library(dplyr)
library(ggplot2)
theme_set(theme_bw(20))
df <- df0 %>% filter(publication_year >=1640 & publication_year <=1828) %>%
              group_by(catalog, publication_decade) %>% tally()
p <- ggplot(df, aes(x = publication_decade, y = n, color = catalog)) + 
       geom_point() + 
       geom_line() + 
       ylab("Documents (n)") + xlab("Publication year") +
       scale_color_manual(values = c("blue", "darkgreen")) +
       guides(color = guide_legend(reverse = TRUE, title = "")) #+ 
print(p)

```

---

### Rise of the octavo sized book as vehicle of Enlightenment

Paper consumption: document formats over time

```{r Gaudeamus-3, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=13, dev="CairoPDF"}
theme_set(theme_bw(22))
pics <- list()
for (catal in unique(df0$catalog)) {
  df2 <- df0 %>% filter(catalog == catal) %>% 
                 group_by(publication_decade, gatherings) %>% 
		 summarise(paper2 = sum(paper, na.rm = TRUE)/1e6, n = n()) 
  df2$paper <- df2$paper2; df2$paper2 <- NULL # Rename
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  df2$highlight <- rep("Other", nrow(df2))
  df2$highlight[df2$gatherings == "8vo"] <- "Octavo"

  p <- ggplot(df2, aes(y = paper,
                       x = publication_decade,
		       shape = gatherings
		       ))		       
  p <- p + geom_point(size = 3)
  p <- p + geom_smooth(method = "loess", size = 1,
             aes(color = highlight, 
	          fill = highlight, 
		  linetype = gatherings), color = "black")
  p <- p + scale_color_manual(values = c("black", "lightgray"))
  p <- p + scale_fill_manual(values = c("black", "lightgray"))  	 
  p <- p + xlab("Vuosi")
  p <- p + ylab("Paperinkulutus (miljoonia arkkeja)")
  p <- p + guides(linetype = guide_legend(keywidth = 5, title = "Muoto"),
       	             shape = guide_legend(keywidth = 5, title = "Muoto"),
		     fill = guide_legend(title = "Korostus"),
		     color = guide_legend(title = "Korostus")
		     )
  p <- p + ylim(c(0, max(df2$paper, na.rm = TRUE)))
  p <- p + ggtitle(catal)
  p <- p + theme(plot.title = element_text(hjust = 0))
  pics[[catal]] <- p
}

library(cowplot)
# https://cran.r-project.org/web/packages/cowplot/vignettes/shared_legends.html

pics[[1]] <- pics[[1]] + xlab("")
prow <- plot_grid( 
           pics[[1]] + theme(legend.position="none"),
           pics[[2]] + theme(legend.position="none"),
           align = 'vh',
           #labels = c("Fennica", "Kungliga"),
           hjust = -1.1,
           vjust = 3,
	   label_size = 18,
           nrow = 2
           )
# extract the legend from one of the plots
# (clearly the whole thing only makes sense if all plots
# have the same legend, so we can arbitrarily pick one.)
legend <- get_legend(pics[[1]])

# add the legend to the row we made earlier. Give it one-third of the width
# of one plot (via rel_widths).
p <- plot_grid( prow, legend, rel_widths = c(3, .6))
print(p)
```

```{r Gaudeamus-Vaasa, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=6, dev="CairoPDF"}
library(dplyr)
theme_set(theme_bw(22))

df2 <- df.combined.preprocessed %>% filter(!is.na(publication_decade)) %>% 
       	       filter(catalog == "Fennica") %>%
       	       filter(publication_year >= 1760) %>%
       	       filter(publication_year <= 1919) %>% 
               filter(publication_place == "Vaasa") %>% 
	       group_by(publication_decade) %>% 
	       tally() %>% 
	       arrange(publication_decade)

p <- ggplot(df2, aes(y = n, x = publication_decade))		       
   p <- p + geom_bar(stat = "identity")
   p <- p + xlab("Vuosi")
   p <- p + ylab("Dokumenttia (n)")
   # Shift the bars so that each bar starts from the first year of that decade
   # for clarity. 
   p <- p + scale_x_continuous(breaks = c(1764,1809,1863,1917) - 4,
     	                       labels = c(1764,1809,1863,1917))
print(p)
```

```{r Gaudeamus-Turku, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=6, dev="CairoPDF"}
library(dplyr)
theme_set(theme_bw(22))
df2 <- df.combined.preprocessed %>% filter(!is.na(publication_decade)) %>% 
       	       filter(catalog == "Fennica") %>%
       	       filter(publication_year >= 1640) %>%
       	       filter(publication_year <= 1919) %>% 	       
               filter(publication_place == "Turku") %>% 
	       group_by(publication_decade) %>% 
	       tally() %>% 
	       arrange(publication_decade)

p <- ggplot(df2, aes(y = n, x = publication_decade))		       
   p <- p + geom_bar(stat = "identity")
   p <- p + xlab("Vuosi")
   p <- p + ylab("Dokumenttia (n)")
   # Shift the bars so that each bar starts from the first year of that decade
   # for clarity. 
   p <- p + scale_x_continuous(breaks = c(1640,1764,1809,1863, 1917) - 4,
     	                       labels = c(1640,1764,1809,1863, 1917))
print(p)
```

```{r Gaudeamus-Helsinki, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=6, dev="CairoPDF"}
library(dplyr)
theme_set(theme_bw(22))
df2 <- df.combined.preprocessed %>% filter(!is.na(publication_decade)) %>% 
       	       filter(catalog == "Fennica") %>%
       	       filter(publication_year >= 1740) %>%
       	       filter(publication_year <= 1919) %>% 	       
               filter(publication_place == "Helsinki") %>% 
	       group_by(publication_decade) %>% 
	       tally() %>% 
	       arrange(publication_decade)

p <- ggplot(df2, aes(y = n, x = publication_decade))		       
   p <- p + geom_bar(stat = "identity")
   p <- p + xlab("Vuosi")
   p <- p + ylab("Dokumenttia (n)")
   # Shift the bars so that each bar starts from the first year of that decade
   # for clarity. 
   p <- p + scale_x_continuous(breaks = c(1764,1809,1863, 1917) - 4,
     	                       labels = c(1764,1809,1863, 1917))
print(p)
```

```{r Gaudeamus-Suomi, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=6, dev="CairoPDF"}
library(dplyr)
theme_set(theme_bw(22))
df2 <- df.combined.preprocessed %>% filter(!is.na(publication_decade)) %>% 
       	       filter(catalog == "Fennica") %>%
       	       filter(country == "Finland") %>%	       
       	       filter(publication_year >= 1640) %>%
       	       filter(publication_year <= 1919) %>% 	       
	       group_by(publication_decade) %>% 
	       tally() %>% 
	       arrange(publication_decade)

p <- ggplot(df2, aes(y = n, x = publication_decade))		       
   p <- p + geom_bar(stat = "identity")
   p <- p + xlab("Vuosi")
   p <- p + ylab("Dokumenttia (n)")
   # Shift the bars so that each bar starts from the first year of that decade
   # for clarity. 
   p <- p + scale_x_continuous(breaks = c(1640,1764,1809,1863, 1917) - 4,
     	                       labels = c(1640,1764,1809,1863, 1917))
print(p)
```

```{r Gaudeamus-Suomi2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=6, dev="CairoPDF"}
library(dplyr)
theme_set(theme_bw(22))
df2 <- df.combined.preprocessed
df2$publication_place[!df2$publication_place %in% c("Helsinki", "Turku", "Vaasa")] <- "Muu"
df2 <- df2 %>%
       	       filter(!is.na(publication_decade)) %>% 
       	       filter(catalog == "Fennica") %>%
       	       filter(country == "Finland") %>%
       	       filter(publication_year >= 1640) %>%
       	       filter(publication_year <= 1919) 
p <- ggplot(df2, aes(x = publication_decade)) +
   geom_bar(position = "stack", aes(y = ..count.., fill = publication_place), color = "black") +
   scale_fill_grey() +   
   xlab("Vuosi") +
   ylab("Dokumenttia (n)") +
   scale_x_continuous(breaks = c(1640, 1764, 1809, 1863, 1917),
     	              labels = c(1640, 1764, 1809, 1863, 1917)) +
   guides(fill = guide_legend(keywidth=1, title = "Julkaisupaikka")) +
   theme(legend.position = c(0.16, .7)) 
print(p)

library(tidyr)
table_hs <- df2 %>% group_by(publication_decade, publication_place) %>%
                    tally() %>%
		    spread(publication_place, n, fill = 0)
library(openxlsx)		    
# openxlsx::write.xlsx(table_hs, file = "Table_HS_20180107.xlsx")
```