### 2018-09-11 inst/examples/201808-HistoricalMethods-Supplement.Rmd

---
title: "A Quantitative Approach to Book Printing in Sweden and Finland, 1640–1828: supplementary material"
author: "Mikko Tolonen, Leo Lahti, Hege Roivainen, Jani Marjanen (2018)"
output: word_document
---

This supplementary document provides a summary of the data fields
analyzed in the main text.

```{r init, echo=FALSE, message=FALSE, warning=FALSE}
library(devtools)
#install_github("haozhu233/kableExtra")
library(kableExtra)
load_all("~/Rpackages/bibliographica")
load_all("~/Rpackages/fennica")
library(skimr)
library(ggplot2)
theme_set(theme_bw(15))
library(tidyverse)

# To be run after all historical methods analyses
# df.combined.preprocessed <- readRDS("data/unified/df.combined.Rds")
df0 <- df.full

d <- list()

fen <- subset(df0, Catalogue == "FNB")
kun <- subset(df0, Catalogue == "SNB")
cats <- list(FNB = fen, SNB = kun)

selected.fields <- c("author", "publication_year", "publication_place", "gatherings", "title", "pagecount", "publisher", "language_primary", "Topic")

for (cat in unique(df0$Catalogue)) {

  dd <- subset(df0, Catalogue == cat)[, selected.fields]

  complete <- round(100 * colMeans(!is.na(dd)), 1)
  unique <- apply(dd, 2, function (x) {length(unique(x))})

  d[[cat]] <- data.frame(Field = colnames(dd), Complete = complete, Unique = unique)

}
```

### Data availability

Records in the final data set used in the analyses:

 * FNB: ```r nrow(subset(df0, Catalogue == "FNB"))```
 * SNB: ```r nrow(subset(df0, Catalogue == "SNB"))```

Data availability for the key fields analyzed in this work (%).

```{r complete, echo=FALSE, message=FALSE, warning=FALSE}
w <- data.frame(Field = map_fieldnames(colnames(dd)),
                "FNB" = d$FNB$Complete,
	        "SNB" = d$SNB$Complete) %>%
	     arrange(desc(FNB))

colnames(w) <- c("Field", "FNB (%)", "SNB (%)")
# kable(w)

# Add Footnote
w[which(w$Field == "Topic"), "SNB (%)"] <- paste0(as.character(w[which(w$Field == "Topic"), "SNB (%)"]), "*")
w[which(w$Field == "Publisher"), "SNB (%)"] <- paste0(as.character(w[which(w$Field == "Publisher"), "SNB (%)"]), "*")

kable <- knitr::kable
knitr::kable(w, format = "markdown")

#%>% kable_styling(latex_options = c("striped", "hold_position"))

#kable(w, format = "latex", booktabs = T, caption = "Demo Table") 
#  kable_styling(latex_options = c("striped", "hold_position"),
#                full_width = F) %>%
#  
```

$^*$The SNB publishers and topics are not used due to the low coverage of harmonized data entries.

### Authors

Number of unique authors and all documents with author information:

```{r author, echo=FALSE, message=FALSE, warning=FALSE}
field <- "author"
df <- df0
df$field <- df[[field]]
df <- df %>% group_by(Catalogue) %>%
              summarise(Unique = length(unique(field)),
	                All = sum(!is.na(field)))
colnames(df) <- gsub("Catalogue", "Catalogue", colnames(df))
knitr::kable(df, digits = 1, format = "markdown")
```

The most common authors in FNB:

```{r authors_fen, echo=FALSE, message=FALSE, warning=FALSE}
tab <- top(fen, field, 10, "data.frame", round = 1)
colnames(tab) <- gsub("author", "Author", colnames(tab))
knitr::kable(tab, format = "markdown")
```

The most common authors in SNB:

```{r authors_kun, echo=FALSE, message=FALSE, warning=FALSE}
tab <- top(kun, field, 10, "data.frame", round = 1)
colnames(tab) <- gsub("author", "Author", colnames(tab))
knitr::kable(tab, format = "markdown")
```

### Languages

In FNB, we have identified ```r sum(subset(df0, Catalogue == "FNB")$multilingual)``` multilingual documents in the investigated time period. In SNB, the multilingual documents have not been analyzed yet. We use the primary language of the document for summaries.

Number of identified primary languages and all documents with language information.

```{r languages , echo=FALSE, message=FALSE, warning=FALSE}
field <- "language_primary"
df <- df0
df$field <- df[[field]]
df <- df %>% group_by(Catalogue) %>%
              summarise(Unique = length(unique(field)),
	                All = sum(!is.na(field)))
colnames(df) <- gsub("Catalogue", "Catalogue", colnames(df))
colnames(df) <- gsub("language", "Language", colnames(df))
knitr::kable(df, format = "markdown")
```

The most common (primary) languages in FNB:

```{r languages_fen, echo=FALSE, message=FALSE, warning=FALSE}
tab <- top(fen, field, 10, "data.frame", round = 1)
colnames(tab) <- gsub("language_primary", map("language_primary", fieldname_table(), from = "Data", to = "Name"), colnames(tab))
knitr::kable(tab, format = "markdown")
```

The most common (primary) languages in SNB:

```{r languages_kun, echo=FALSE, message=FALSE, warning=FALSE}
tab <- top(kun, field, 10, "data.frame", round = 1)
colnames(tab) <- gsub("language_primary", map("language_primary", fieldname_table(), from = "Data", to = "Name"), colnames(tab))
knitr::kable(tab, format = "markdown")
```

### Formats (Gatherings)

Proportion (%) of gatherings in the FNB and SNB Catalogues. The gatherings with a prevalence less than 1% in both Catalogues have been combined into the "Other" category.

```{r gatherings, echo=FALSE, message=FALSE, warning=FALSE}
df <- df0 %>% group_by(Catalogue, gatherings) %>%
              summarise(n = n())
df <- spread(df, Catalogue, n, fill = 0)
colnames(df) <- c("Gatherings", "FNB", "SNB")
df$Gatherings <- as.character(df$Gatherings)
df$Gatherings <- gsub("NA", "Missing", df$Gatherings)

df2 <- df
df2$FNB <- round(100 * df2$FNB / nrow(fen), 1)
df2$SNB <- round(100 * df2$SNB / nrow(kun), 1)

# Combine gatherings with <1% prevalence into "Other category"
inds <- which(rowMeans(apply(df2[, 2:3], 2, as.numeric) < 1) == 1)
df2 <- rbind(df2, c("Other", colSums(apply(df2[which(rowMeans(df2[, 2:3] < 1) == 1),], 2, as.numeric))[-1])) %>% slice(-inds)
df2$Gatherings <- gsub("NA", "Missing", df2$Gatherings)
df2$Gatherings[is.na(df2$Gatherings)] <- "Missing"
accepted.gatherings <- df2$Gatherings

names(df2) <- gsub("NB", "NB (%)", names(df2))
df2$Gatherings2 <- map_gatherings(unlist(df2[,1]), "Short", "Name")
df2$Gatherings <- apply(cbind(df2$Gatherings2, paste0("(", unlist(df2$Gatherings), ")")), 1, function (x) {paste(x, collapse = " ")})
df2$Gatherings2 <- NULL
df2$Gatherings <- gsub(" \\(Missing\\)", "", df2$Gatherings)
df2$Gatherings <- gsub(" \\(Other\\)", "", df2$Gatherings)
knitr::kable(df2, format = "markdown")

# Take the changes also in original data now for later figs
inds <- which(!df0$gatherings %in% accepted.gatherings)
df0$gatherings[inds] <- "Other"
```

### Pages

Page count variation per gatherings.

```{r pagecount, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, figh.height=1}
theme_set(theme_bw(15))
df <- df0
df <- df[which(!df$gatherings == "NA"), ]
df <- subset(df, !is.na(pagecount))
df$gatherings <- order_gatherings(df$gatherings)

cols <- default_colors("gatherings", map_gatherings(levels(df$gatherings), from = "Short", to = "Name"))

p <- ggplot(df) +
       geom_histogram(aes(x = pagecount,
		    group = gatherings,       
                    #y = ..count../sum(..count..),
                    fill = gatherings),
                    color = "black",
		    alpha = 0.5
		    ) + 
       labs(x = "Page count (n)",
            y = "Frequency (n)") +
       scale_x_log10(breaks = c(1, 10, 100, 1000)) +
       scale_y_log10(breaks = c(1, 10, 100, 1000)) +
       scale_fill_manual(values = unname(cols)) +       
       facet_grid(Catalogue ~ gatherings) +
       guides(fill = "none") + 
       labs(y = "Frequency (n)") +
       theme(axis.text = element_text(size=12))
print(p)
```

### Publication places

Number of unique publication places and all documents with publication place information.

```{r places, echo=FALSE, message=FALSE, warning=FALSE}
field <- "publication_place"
df <- df0
df$field <- df[[field]]

df <- df %>% group_by(Catalogue) %>%
              summarise(Unique = length(unique(field)),
	                All = sum(!is.na(field)))
colnames(df) <- gsub("Catalogue", "Catalogue", colnames(df))
knitr::kable(df, format = "markdown")
```

The most common publication places in FNB:

```{r places_fen, echo=FALSE, message=FALSE, warning=FALSE}
tab <- top(fen, field, 10, "data.frame", round = 1)
colnames(tab) <- gsub("publication_place", "Publication place", colnames(tab))
knitr::kable(tab, format = "markdown")
```

The most common publication places in SNB:

```{r places_kun, echo=FALSE, message=FALSE, warning=FALSE}
tab <- top(kun, field, 10, "data.frame", round = 1)
colnames(tab) <- gsub("publication_place", "Publication place", colnames(tab))
knitr::kable(tab, format = "markdown")
```

### Publication years

Title count per decade for FNB and SNB.

```{r year, echo=FALSE, message=FALSE, warning=FALSE, fig.width=9, figh.height=4, out.width = "300px"}
theme_set(theme_bw(20))
df <- df0
names(df) <- gsub("Catalogue", "Catalogue", names(df))
df <- df %>% group_by(Catalogue, publication_decade) %>% tally()
p <- ggplot(df, aes(x = publication_decade, y = n, fill = Catalogue)) +
       geom_bar(color = "black",
         position = "dodge", stat = "identity") +
       guides(fill = guide_legend(reverse = TRUE)) +
       labs(x = "Publication decade",
            y = "Frequency (n)") +
       scale_fill_manual(values = rev(c("darkgray", "black"))) #+
       #theme(axis.text = element_text(size=15))
print(p)
```

### Publishers

In FNB, the data includes ```r  length(unique(subset(df0, Catalogue == "FNB")$publisher))``` unique publishers, and ```r  round(100 * mean(!is.na(subset(df0, Catalogue == "FNB")$publisher), na.rm = TRUE))```% of the documents have publisher information. We have not analyzed publishers in SNB, where the availability of harmonized publisher information is currently low (```r  round(100 * mean(!is.na(subset(df0, Catalogue == "SNB")$publisher), na.rm = TRUE))```%).

The most common publishers in FNB (the total count and percentage of all documents are shown):

```{r publishers_fen, echo=FALSE, message=FALSE, warning=FALSE}
tab <- top(fen, "publisher", 10, "data.frame", round = 1)
colnames(tab) <- gsub("publisher", "Publisher", colnames(tab))
knitr::kable(tab, format = "markdown")
```

### Topics

We use here the term "Topic" for the "Topical term" MARC field 650a for brevity. In FNB, the data includes ```r  length(unique(subset(df0, Catalogue == "FNB")$Topic))``` unique topics. ```r  round(100 * mean(!is.na(subset(df0, Catalogue == "FNB")$Topic), na.rm = TRUE))```% of the documents have topic information. We have not analyzed topics in SNB, where the availability of harmonized topics is currently low (```r  round(100 * mean(!is.na(subset(df0, Catalogue == "SNB")$Topic), na.rm = TRUE))```%).

The most common topics in FNB (English translation in parentheses):

```{r topics_fen, echo=FALSE, message=FALSE, warning=FALSE}
tab <- top(fen, "Topic", 10, "data.frame", round = 1)
colnames(tab) <- gsub("Topic", "Topic", colnames(tab))

trans = c("hautajaiset" = "funerals",
"etsintäkuulutukset" = "warrants",
"häät" = "weddings",
"yliopistot;juhlamenot;Turku" = "universities;ceremonies;Turku",
"arkkiveisut;virret" = "broadside ballads;hymns",
"dogmatiikka" = "dogmatics",
"moraali" = "moral philosophy",
"homiletiikka" = "homiletics",
"filosofia" = "philosophy",
"eksegetiikka" = "exegetics")
tab$Topic <- paste(tab$Topic, paste0("(", unname(trans[tab$Topic]), ")"))
knitr::kable(tab, format = "markdown")
```

### MARC fields

The following table indicates the MARC fields for each data type in
our harmonized data collection. The language information was picked from the
MARC fields 008lang and 041a for SNB and FNB, respectively.

```{r fieldtable, echo=FALSE, message=FALSE, warning=FALSE}
f <- field_table()

fennica_fields <- c(
"041a",
"041h", 
"100a",
"100d",
"240a",
"245a",
"245b",
"260a",
"260b",
"260c",
"300a",
"300b",
"300c",
"300e",
"310a",
"362a",
"500a",
"502a",
"502c",
"502d",
"510a",
"510c",
"650a",
"651a",
"710a",
"720a",
"785t",
"852a")

kungliga_fields <- c(
setdiff(fennica_fields, c("041a", "041h")),
"008lang",
"100a",
"100d",
"110a",
"240a",
"245a",
"245b",
"245c",
"260a",
"260b",
"260c",
"260e",
"260f",
"300a",
"300b",
"300c",
"300e",
"310a",
"362a",
"440a",
"440v",
"500a",
"502a",
"502c",
"502d",
"510a",
"510c",
"650a",
"650x",
"650y",
"650z",
"651a",
"700a",
"700d",
"710a",
"720a",
"740a",
"772c",
"772d",
"772t",
"785t",
"852a",
"852j",
"852z",
"866x",
"900a",
"900d",
"900u",
"976a",
"976b")

all_fields <- union(fennica_fields, kungliga_fields)

df <- data.frame(MARC = all_fields)
df$Description <- f[match(df$MARC, f$field), "concept"]
df$SNB  <- c("", "x")[1 + df$MARC %in% kungliga_fields]
df$FNB  <- c("", "x")[1 + df$MARC %in% fennica_fields]
df <- df %>% arrange(MARC)

knitr::kable(df, format = "markdown")
```