### 2018-12-11 inst/examples/swedish_gap.Rmd
Swedish gap analysis

---
title: "Gap in Swedish"
author: "Helsinki Computational History Group / Leo"
date: "`r date()`"
---

# Fennica summary

```{r init, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=15}
library(stringr)
library(dplyr)
library(reshape2)
library(ggplot2)
ntop <- 20

# Read the preprocessed data
df <- readRDS("data/unified/polished/df.Rds")
```

```{r gap, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=15}
theme_set(theme_bw(20))
dfs <- df %>% filter(language_primary == "Swedish") %>%
       	      group_by(publication_decade) %>%
              tally()
p <- ggplot(dfs, aes(x = publication_decade, y = n)) +
       #geom_point() +
       #geom_smooth() +
       geom_bar(stat = "identity") +
       labs(x = "Publication decade",
            y = "Title count (n)",
	    title = "Title count for Swedish publications (primary language)"
	    )
print(p)
```