### 2021-01-12 inst/examples/language.Rmd

---
title: "Language summaries"
author: "Leo Lahti"
date: "`r Sys.Date()`"
output: markdown_document
---

## Language

 * `r length(unique(unlist(strsplit(as.character(df$languages), ";"))))-1` [unique languages](output.tables/language_accepted.csv)
 * `r length(unique(df$language_primary))-1` [unique primary languages](output.tables/language_accepted.csv)  
 * `r sum(!df$multilingual, na.rm = TRUE)` single-language documents (`r round(100 * mean(!df$multilingual, na.omit = TRUE), 2)`%)
 * `r sum(df$multilingual, na.rm = TRUE)` multilingual documents (`r round(100 * mean(df$multilingual, na.omit = TRUE), 2)`%) 
 * [Conversions from raw to preprocessed language entries](output.tables/language_conversions.csv) 
 * `r sum(is.na(df$languages) | df$language_primary == "Undetermined")` documents (`r round(100*mean(is.na(df$languages) | df$language_primary == "Undetermined"), 2)`%) with [unrecognized language](output.tables/language_discarded.csv)

Language codes are from [MARC](http://www.loc.gov/marc/languages/language_code.html); new custom abbreviations can be added in [this table](https://github.com/COMHIS/fennica/blob/master/inst/extdata/language_abbreviations.csv).

Title count per language (including multi-language documents; note the log10 scale):

```{r summarylang, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=5}
top_plot(df, "languages", ntop = ntop, log10 = TRUE) 
```

### Top languages

Number of documents assigned with each language (top-10). For a complete list,
see [accepted languages](output.tables/language_accepted.csv).

```{r summarylang2, echo=FALSE, results='asis'}
x <- top(df, "languages")
tab <- cbind(names(x), unname(x), round(100 * unname(x/nrow(df)), 1))
colnames(tab) <- c("Language", "Documents (n)", "Fraction (%)")
kable(head(tab, 10))
```