### 2021-01-29 inst/examples/title.Rmd
updated title page

---
title: "Title field: summary"
author: "Leo Lahti"
date: "`r Sys.Date()`"
output: md_document
---

### Titles

 * [Unique accepted titles](output.tables/title_accepted.csv): `r length(unique(df$title))`
 * [Unique discarded titles in original data](output.tables/title_discarded.csv) (excluding NA cases): `r as.vector(countLines(paste0(output.folder, "/title_discarded.csv")) - 1)` 
 * Original documents with non-NA titles: `r sum(!is.na(df$title))` / `r nrow(df)` (`r round(100*mean(!is.na(df$title)), 1)`%)
 * Original documents with missing (NA) titles `r sum(is.na(df$title))` / `r nrow(df)` documents (`r round(100*mean(is.na(df$title)), 1)`%)

Top-`r ntop` titles and their title counts.

```{r summarytitle, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
library(ggplot2)
p <- top_plot(df, "title", ntop = ntop) +
       scale_y_log10() +
       labs(y = "Entries (n)", title = "Top titles")
print(p)
```

```{r uniquetitles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
tab <- table(df$title)
d <- data.frame(title = names(tab), n = as.numeric(tab))

library(ggplot2)
p <- ggplot(d, aes(x = n)) +
       geom_histogram() +
       scale_x_log10() +
       scale_y_log10() +
       labs(x = "Title duplicates (n)", y = "Documents (n)")
print(p)
```