### 2021-02-17 inst/examples/overview.Rmd
moved many functions from bibliographica to fennica for simplicity and maintenance

---
title: "Preprocessing overview"
author: "`r author`"
date: "`r Sys.Date()`"
output: markdown_document
---

# Fennica preprocessing summary

The original data set contains `r nrow(df.orig)` documents.

The following fields have been harmonized.

  * [Author info](author.md)
  * [Gender info](gender.md)
  * [Publisher info](publisher.md)
  * [Publication geography](publication_place.md)
  * [Publication year info](publicationyear.md)
  * [Titles](title.md)  
  * [Page counts](pagecount.md)
  * [Physical dimension](dimension.md)    
  * [Document and subject topics](topic.md)
  * [Languages](language.md)
  * [Corporate](corporate.md)
  * [Holder](holder.md)
  * [Note_510c](note_510c.md)
  * [Note_dissertation](note_dissertation.md)
  * [Note_granter](note_granter.md)
  * [Note_source](note_source.md)
  * [Subject_topic](subject_topic.md)
  * [Successor](successor.md)
  * [Title_uniform](title_uniform.md)
  * [Uncontrolled](uncontrolled.md)
See also [further preprocessing details](details.Rmd).