### 2021-02-18 inst/examples/publication_frequency.Rmd
ok

---
title: "Publication frequency summaries"
author: "Leo Lahti"
date: "`r Sys.Date()`"
output: markdown_document
---

### Publication frequency

Publication frequency information is available for `r sum(!is.na(df$publication_frequency_text))` documents (`r round(100*mean(!is.na(df$publication_frequency_text)))`%). The links are invalid if the lists are empty. The (estimated annual) frequencies are converted to plain text according to their closest match in [this table](https://github.com/COMHIS/fennica/blob/master/inst/extdata/frequency_conversions.csv).

[Publication frequency accepted](output.tables/publication_frequency_accepted.csv)

[Publication frequency conversions](output.tables/publication_frequency_conversion.csv)

[Publication frequency discarded](output.tables/publication_frequency_discarded.csv)