### 2021-02-18 inst/examples/publication_interval.Rmd

---
title: "Publication interval summaries"
author: "Leo Lahti"
date: "`r Sys.Date()`"
output: markdown_document
---

### Publication interval

Publication interval is available for `r sum(!is.na(df$publication_interval_from) | !is.na(df$publication_interval_till))` documents (`r round(100*mean(!is.na(df$publication_interval_from) | !is.na(df$publication_interval_till)))`%).

[Publication interval accepted](output.tables/publication_interval_accepted.csv)

[Publication interval conversions](output.tables/publication_interval_conversion_nontrivial.csv)

[Publication interval discarded](output.tables/publication_interval_discarded.csv)