### 2021-03-03 inst/examples/author_name.Rmd
check

---
title: "author_name"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path=paste0("figure/rmd_", field, "_"))
```

* [Unique accepted entries](`r file_accepted`): `r length(unique(df[[field]]))`
* [Unique discarded entries in original data](`r file_discarded`) (excluding NA cases): `r as.vector(countLines(file_discarded) - 1)`
* Original documents with non-NA titles: `r sum(!is.na(df[[field]]))` / `r nrow(df)` (`r round(100*mean(!is.na(df[[field]])), 1)`%)
* Original documents with missing (NA) titles `r sum(is.na(df[[field]]))` / `r nrow(df)` documents (`r round(100*mean(is.na(df[[field]])), 1)`%)

 Top-`r ntop` titles and their title counts.

```{r summarytitle, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
library(ggplot2)
theme_set(theme_bw(base_size = 12))
p <- top_plot(df, "author_name", ntop = ntop, log10 = TRUE)+
  labs(y = "Entries (n)", title = "Top titles")
print(p)
```

Frequency of unique titles:

```{r uniquetitles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
tab <- table(df$author_name)
d <- data.frame(title = names(tab), n = as.numeric(tab))

library(ggplot2)
theme_set(theme_bw(base_size = 12))
p <- ggplot(d, aes(x = n)) +
  geom_histogram(color="black", fill="white") +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Title duplicates (n)", y = "Documents (n)")
print(p)
```

## Authors

* `r length(unique(df$author_name))` [unique authors](output.tables/author_accepted.csv) These final names capture all name variants from the custom [author synonyme table](https://github.com/COMHIS/fennica/blob/master/inst/extdata/ambiguous-authors.csv), and exclude known pseudonymes (see below). If multiple names for the same author are still observed on this list, they should be added on the [author synonyme table](https://github.com/COMHIS/fennica/blob/master/inst/extdata/ambiguous-authors.csv).
 * `r sum(!is.na(df$author_name))` documents have unambiguous author information (`r round(100*mean(!is.na(df$author_name)))`%). 
 * `r length(unique(subset(df.preprocessed, author_pseudonyme)$author_name))` [unique pseudonymes](output.tables/pseudonyme_accepted.csv) are recognized based on [custom pseudonyme lists](https://github.com/COMHIS/fennica/blob/master/inst/extdata/names/pseudonymes/custom_pseudonymes.csv).
 * `r length(readLines(paste0(this.folder, "/output.tables/author_discarded.csv")))-2` [discarded author names](output.tables/author_discarded.csv) This list should not include any real authors (if it does, please send a note to the admin). The stopword lists are considered when discarding names.
 * [Author name conversions](output.tables/author_conversion_nontrivial.csv) Non-trivial conversions from the original raw data to final names.

### Auxiliary files

 * [Author synonyme table](https://github.com/COMHIS/fennica/blob/master/inst/extdata/ambiguous-authors.csv)
 * [Stopwords for names](https://github.com/COMHIS/fennica/blob/master/inst/extdata/stopwords_for_names.csv)
 * [Stopwords for titles](https://github.com/COMHIS/fennica/blob/master/inst/extdata/stopwords_titles.csv)
 * [Custom pseudonyme lists](https://github.com/COMHIS/fennica/blob/master/inst/extdata/names/pseudonymes/custom_pseudonymes.csv)

Top-`r ntop` uniquely identified authors and their productivity (title count).

```{r summaryauthorsnames, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5, out.width='430px', fig.show='hold'}
knitr::opts_chunk$set(fig.path = "figure_slides/", dev="CairoPNG")
theme_set(theme_bw(base_size = 12))
p <- NULL
p <- top_plot(df, "author_name", ntop) +
     		  ggtitle(paste("Top authors' names")) +
		  ylab("Documents")
print(p)
```

## summary discraded plot does not work 
```{r summarydiscardedauthors, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5, out.width='430px', fig.show='hold'}
dff = df %>%
      filter(is.na(author) & !is.na(author_name)) %>%
      select(author_name)
p <- top_plot(dff,  "author_name", ntop) + ggtitle("Top discarded authors")
print(p)
```

### Author productivity

Title count versus paper consumption (all authors):

```{r authortitlespapers, fig.height=8, fig.width=18, echo=FALSE, warning=FALSE}
theme_set(theme_bw(base_size = 12))
res1 <- compare_title_paper(df, "author_name", plot.mode = "text")
res2 <- compare_title_paper(df, "author_name", plot.mode = "point")
grid.arrange(res2$plot, res1$plot, nrow = 1)
#kable(res1$table)
```

#Publication_decade not found
#```{r summaryTop10authorstimeline, fig.height=8, fig.width=16, echo=FALSE}
theme_set(theme_bw(20))
top.authors <- names(top_n(df, field = "author_name", n = 3))
dfs <- df %>% filter(author %in% top.authors) %>%
     	 group_by(author_name, publication_decade) %>%
     	 tally() %>%
     	 arrange(publication_decade)
p <- ggplot(dfs, aes(x = publication_decade, y = n, fill = author)) +
       geom_bar(stat = "identity", position = "stack", color = "black") +
       xlab("Publication Decade") +
       ylab("Title Count") +
       scale_fill_grey() +
       guides(fill = guide_legend("Author")) +
       ggtitle("Title count timeline for the top authors")
print(p)
##```

###Publication year not found
##```{r topauth, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=13, fig.height=6}
df <- df.preprocessed
top.authors <- names(rev(sort(table(df.preprocessed$author))))[1:10]
df2 <- df %>% filter(author %in% top.authors) %>% group_by(publication_year, author) %>% summarize(paper = sum(paper, na.rm = TRUE), n = n())

p <- ggplot(df2, aes(x = publication_year, y = paper, group = author, color = author))
p <- p + geom_point()
p <- p + geom_line()
p <- p + ggtitle("Paper consumption per author")
p <- p + xlab("Year")
p <- p + scale_y_log10()
p <- p + ylab("Paper consumption")
print(p)

p <- ggplot(df2, aes(x = publication_year, y = n, group = author, color = author))
p <- p + geom_point()
p <- p + geom_line()
p <- p + ggtitle("Documents per author")
p <- p + xlab("Year")
p <- p + scale_y_log10()
p <- p + ylab("Documents (n)")
print(p)
###```