### 2021-03-04 inst/examples/publisher.Rmd
change

---
title: "Publisher preprocessing summary"
author: "`r author`"
date: "`r Sys.Date()`"
output: md_document
---

### Publishers

 * `r length(unique(df$publisher))` [unique publishers](output.tables/publisher_accepted.csv)

 * `r sum(!is.na(df$publisher))` documents have unambiguous publisher information (`r round(100*mean(!is.na(df$publisher)), 1)`%). This includes documents identified as self-published; the author name is used as the publisher in those cases (if known).

 * [Discarded publisher entries](output.tables/publisher_discarded.csv)

 * [Conversions from original to final names](output.tables/publisher_conversion_nontrivial.csv) (only non-trivial conversions shown)

The `r ntop` most common publishers are shown with the number of documents.

```{r summarypublisher2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=9}
p <- top_plot(df, "publisher", ntop)
p <- p + ggtitle(paste("Top publishers"))
p <- p + scale_y_log10()
p <- p + ylab("Documents")
print(p)
```

Title count versus paper consumption (top publishers):

```{r publishertitlespapers, fig.height=8, fig.width=8, echo=FALSE, warning=FALSE}
#res <- compare_title_paper(df, "publisher", selected = tops)
tops <- names(top(df, field = "publisher", n = 10))
res <- compare_title_paper(df, "publisher")
print(res$plot)  
kable(subset(res$table, publisher %in% tops), digit = 2)
```