### 2021-03-05 inst/examples/publication_time.Rmd
update

---
title: "Publication year summaries"
author: "Leo Lahti"
date: "`r Sys.Date()`"
output: md_document
---

### Publication time (publication year)

[Publication year conversions](output.tables/publication_year_conversion.csv)

[Publication year discarded](output.tables/publication_year_discarded.csv)

Publication year is available for `r sum(!is.na(df$publication_year))` documents (`r round(100*mean(!is.na(df$publication_year)))`%). The publication years span `r paste(range(na.omit(df$publication_year)), collapse = "-")`.

```{r summarypublicationyear, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
# Title count per decade
df2 <- df %>% group_by(publication_decade) %>% filter(publication_decade < 2010) 
p <- ggplot(df2, aes(publication_decade)) +
     geom_bar() + scale_y_log10() +
     ggtitle("Title count timeline")
print(p)
```