### 2022-09-19 inst/examples/index.Rmd

--- 
title: "Harmonized Finnish National Bibliography"
author: "`r author`"
date: "`r Sys.Date()`"
site: "bookdown::bookdown_site"
output:
  bookdown::gitbook: default
documentclass: book
---

In this bookdown project we will disseminate the fennica dataset cleaning process.

```{r source-index, include=FALSE}
source("init.R")
# knit_bookdown <- TRUE in init.R
```