### 2022-11-16 inst/examples/index.Rmd
Bookdown compatibility

--- 
title: "Harmonized Finnish National Bibliography"
author: "Helsinki Computational History Group (COMHIS)"
date: "`r Sys.Date()`"
site: "bookdown::bookdown_site"
output:
  bookdown::gitbook: default
documentclass: book
---

# Preface

In this bookdown project we will disseminate the Finnish National Bibliography (fennica) bibliographic dataset cleaning process.

```{r source-index, include=FALSE}
source("init.R")
# knit_bookdown <- TRUE in init.R
```

This document was generated with the following packages:

```{r system-information}
sessionInfo()
```