### 2022-11-16 inst/examples/publication_interval.Rmd

```{r source-publication-interval, include=FALSE, eval=knit_bookdown}
#publication_interval field
source("publication_interval.R")
```

# Publication interval summaries

Publication interval is available for `r sum(!is.na(df$publication_interval_from) | !is.na(df$publication_interval_till))` documents (`r round(100*mean(!is.na(df$publication_interval_from) | !is.na(df$publication_interval_till)))`%).

[Publication interval accepted](output.tables/publication_interval_accepted.csv)

[Publication interval conversions](output.tables/publication_interval_conversion_nontrivial.csv)

[Publication interval discarded](output.tables/publication_interval_discarded.csv)